### 2018-01-21 paper.tex
first commit

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage{array,multirow}
% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El Paso, Texas USA} 
\acmYear{1997}
%\copyrightyear{2016}

%\acmArticle{4}
%\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}

\begin{document}
%\title{Detecting Emergency Situations Based on Locations in Twitter}
\title{Self-Organized Citizen Sensors for Detecting Emergency Situations in Twitter}
\titlenote{Produces the permission block, and
  copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
  %\texttt{acmart.pdf} document}

\author{Hernan Sarmiento}
%\authornote{Dr.~Trovato insisted his name be first.}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{Deparment of Computer Science, University of Chile}
  \streetaddress{P.O. Box 1212}
   \city{Santiago} 
   \country{Chile}}
\email{hsarmien@dcc.uchile.cl}

\author{Barbara Poblete}
%\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{Deparment of Computer Science, University of Chile}
  \streetaddress{P.O. Box 1212}
   \city{Santiago} 
   \country{Chile}}
\email{bpoblete@dcc.uchile.cl}

\author{Sergio Campos}
%\authornote{This author is the
 % one who did all the really hard work.}
\affiliation{%
  \institution{Department of Geophysics, University of Chile}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Santiago} 
  \country{Chile}}
\email{jaime@dgf.uchile.cl}

% The default list of authors is too long for headers.
%\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
%Most methods to detect emergency situations using Twitter rely on keyword. The problem of keyword-based methods is the need to train in specific domains and language for different type of events, for example: earthquakes, typhoons, terrorist attacks, tornadoes, etc. Our proposal is to rather use the recurring mention of a country locations in microblogging messages to identify such events without using keywords and characterize through of interarrival time probability distributions. Our method use a SVM classifier that is independent of the text language of the messages and can be training and testing among different languages and domains.

Most methods to detect emergency situations using Twitter rely on keyword. The problem of keyword-based methods is the need to train models in specific domains, in multiple languages, and for different types of events, for example: earthquakes, terrorist attacks, floods, etc. In contrast, our proposal is to rather use the recurring references of a location in the metadata of microblogging messages, where users act self-organized citizen sensors in real-locations. In order to identify an emergency situation, we characterize  such events  through frequencies, and probability distributions of the interarrival time of the messages. Our method uses a SVM classifier which is independent of the language of the text, and textual features. Finally, we describe an emergency situation according to geographic impact based on the propagation of the messages in social media between neighboring locations.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10002951.10003227.10003236.10003239</concept_id>
	<concept_desc>Information systems~Data streaming</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Information systems~Data streaming}

\keywords{Emergency Situations, Citizen Sensors, Social Media}

\maketitle

\section{Introduction}
Social media has become a major channel for communication during high-impact real-world events, for example: elections, sports events, emergency situations, etc. In any case, users act as ``citizen sensors'' where they share and post their mood, opinions, photos, videos and geographical point of interest (POI).

During emergency situations, traditional media suffer structural damage and real-time communications could be disrupted. Instead, microblogging has played a critical role in the last fifteen years since it has allowed users share real-time information from people local to the incident as current status, casualties, damages and alerts \cite{kumar2011tweettracker,imran2013extracting,stowe2016identifying,reuterfifteen}. For this reason, researchers have studied the behaviour during these events for to detect, summarize and classifyng messages with the goal to help authorities and the general public with situational awareness to act fast and conscientiously during a crisis situations.

Twitter, a microblogging which allows to share short messages (called \textit{tweets}) and is currently used world-wide by over $300$ million users\footnote{http://www.statista.com/statistics/282087/number-of-monthly-active-twitter-users/}. About $80\%$ of Twitter users access from mobile services which contribute to the inmediacy of diffusion of information specially during a crisis situations \cite{castillo2016big}. One of the main tasks during these situations is to detect a new real-event because as with most social media conversations, messages are often overcomed by irrelevant and redundant noise. In current works \cite{kumar2011tweettracker,caragea2011classifying,ashktorab2014tweedr,imranaidr2014,maldonado2017}, these tasks are solved with methods rely on keywords over \textit{Twitter Public Streaming API}\footnote{https://developer.twitter.com/en/docs/tweets/filter-realtime/overview}. The problem with these keyword-based methods is the need to train in specific domains for different type of events. \cite{olteanu2014} generate set of keywords with different datasets but sometimes arise specific terms for one event, for example \textit{\#eqnz} for Earthquake in New Zealand or \textit{\#pabloph} for Typhoon Pablo in Philippines \cite{potts2011tweeting,bruns2012local,karimi2013classifying}. Furthermore, these set of keywords are trained for a unique language and can not be used in others.

Additionally, several works related to crisis situations show a strong relationship between type of event (what) and the spatiotemporal dimensions (when and where). For example, we can see top trends during an earthquake related to locations \cite{mendoza2010twitter}, strong relationship between proximity to hurricane path and hurricane-related social media activity \cite{kryvasheyeu2016rapid}, extraction of locations and POI during floods \cite{lingad2013location} and mixing geographic information system (GIS) with geo-tagged message to improve disaster maping and real-time event tracking \cite{huang2015disastermapper}.

Based on previous works, the question we address here is: \textit{Is there evidence to detect an emergency situation based on anomaly frequency of messages that contain locations?}.

Here, we propose a method based on recurring references of country locations in message's metadata. For this task, we create a gazetteer tree based on the hierarchy of the specific country, divide the messages into fixed time-windows and compute the frequency, and the probability distributions of the inter-arrival time of the messages for each geographical hierarchy. To detect an emergency situation, we train a SVM classifier by each hierarchy and apply a geographic spread to filter false positives detection considering that an emergency situation can be \textit{Focalized} or \textit{Diffused}.

Our main contribution is to create a methodology that detecting instantaneous emergency situations just using the locations for a specific country. Furthermore, our classifier do not depend of the language to detect a new event since that do not use textual features as input. Moreover, we characterize crisis situations that affect a small or large geographic area.

The paper is organized as follows: we first introduce an overview of relevant literature related to emergency situation and event detection in social media. Next we present a complete description of our proposal divide in four parts. Thereafter, we provide a full description of our dataset and later we summarize our experimental validation and results over the ground truth and online evaluation. Finally, we deliver our discusions, conclusions and future work.

\section{Related Work}
explain first works since 15 years ago
\subsection{Emergency Situation}
\subsubsection{Data Characterization}
crisis lex, information nuggets, twitter under crisis, disaster related messages, practical extrac-
tion of disaster-relevant, Identifying and Categorizing Disaster-Related Tweets,Microblogging
During Two Natural Hazards Events: What Twitter May Contribute to Situational Aware-
ness, Twitter Adoption and Use in Mass Convergence and Emergency Events, Twitter Under
Crisis: Can we trust what we RT?

\subsubsection{Crisis-Related Social Media Monitoring}
tweedr, twicalli, ESA, aidr, twincident, EMERSE, twicalli, tweettracker, twittris

\subsection{Event Detection}
burst detection (jheser), geo-spatial event detection, prediction and characterization of high-
activity events,Emerging event detection in social networks with location sensitivity

\subsection{Location Extraction}
pinpointing, from the justin beiber’s heart, where in the world are you?

\section{Proposed Approach}

In order to provide a complete coverage of location-based detection of emergency situation, we proposed our approach which has four stages. The following information provides details of each stage.

\subsection{Data Pre-Processing}
%In this stage, data is pre-processed from Twitter Streaming API\footnote{algo}, which allows access to subsets equal to $1\%$ of public status descriptions in real-time. With this tool, we can be retrieve messages using a set of keywords or messages from specific locations setting a bounding box. In our approach, we get entire subsets of messages without use keywords or specific locations. Then, we retrieve random messages about any topic and any place in the world.

%To increase the number of messages related to locations, we consider tweets and retweets messages. 
Since that we consider users as citizen sensors, we filter messages depending of native language of each country using the attribute \textit{lang} retrieved from Tweet Metadata\footnote{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object}. For example, if we analyze England, we just consider messages in english. Finally, we remove user mentions, urls,  special characters and apply text tokenization . In this case, we do not remove hashtags or stopwords because some locations can be included in hashtags terms and some places can be contained stopword that can differentiate them from other similar non-related location terms.

\subsection{Signals Creation}
The problems that we address in this stage is how to infer locations from micro-blog messages to track signal over time in the Stream and what is the lowest hierarchy level to find and assign to this signal.

\subsubsection{Geographical Hierarchy}\label{sssec:geohie}
Using the idea of \textit{Gazetteer as a Tree} presented in \cite{yin2014pinpointing} in which each place is associated with a canonical taxonomy node, we construct our gazetteer tree based on Geonames\footnote{http://download.geonames.org/export/dump/} and Wikipedia\footnote{http://www.wikipedia.org/} for each country to analyze. However, in \cite{yin2014pinpointing} the gazetteer hierarchy presents four-levels where the lowest level represents a specific POIs. In our approach, we use a subset of the gazetteer hierarchy with three-levels, in which the lowest level is represented by City since a large amount of users specify their location down to the city scale \cite{hecht2011tweets}. For example, if we have the \textit{city:Manchester}, we associate this location with \textit{region-state:North West} and also with \textit{country:England}. Furthermore, the name of locations are considered just in the native language of the country. For instance, in the case of Italy locations, we consider \textit{Roma} and not \textit{Rome}.

\subsubsection{Finding Locations in Micro-Blog Metadata}

The structure of the Tweet Metadata allows to obtain information about the message and the user who shares a message. Considering the geographical hierarchy explained in the previous section, we search these locations on different levels of metadata and create one-signal for each as following:

\begin{itemize}
	\item Tweet Text: location is mentioned in the attribute \texttt{text} of tweet object, that is, on the body of the message.
	\item User Location: location is mentioned in the attribute \texttt{location} inside the \texttt{user object}, that is, the location set by the user in his profile.
	\item Tweet Text - User Location: location is mentioned in the attribute \texttt{text} of tweet object and also location is mentioned in the attribute \texttt{location} inside the \texttt{user object}. This means that the location is mentioned on the body on the message and the user who share message has the same location in his profile. In this case, tweet text and user location can be different in the smallest hierarchy but in the highest level can be equal. 
\end{itemize}

In this way, mixing geographical hierarchy and locations in micro-blog metadata, we create $NxM$ signals where $N$ is the number of locations obtained by gazetteer tree and $M$ is the number of metadata-level extracted from Tweet Object. In instance, we create a signal for \textit{city:Manchester} and we find this hierarchy in \textit{metadata:Tweet Text} and also in \textit{metadata:User Location}. That means that we track the mention of \textit{city:Manchester} at the level of the body of message and at the level of the location of the user profile individually.

\subsection{Time-Windows}
In this stage we address the problems of how to divide and determine the time-windows size to detect a new emergency situation and what variables by time-window allow it.

\subsubsection{Determining optimal window size}

According to \cite{guzman2013line}: ``If the window size is to small, the ocurrence of empty windows for a term increases, making the noise rate increate and frequency rate tend to zero. On the other hand, if the window size is too large, the stability of the signal becomes constant and busrty keyword detection is delayed''. Using this definition, we divide our signals into windows of six minutes because it divides a 24-hour day exactly, making the analysis easier to understand and to compare from different days.

\subsubsection{Normalized Frequency}
We compute the number of the messages of each time-window by signal. To normalize frequency, we compute \textit{z-score} as following:

\begin{equation} \label{eq:1}
zscore = \frac{x_{i} - \mu_{k} }{\sigma_{k}}
\end{equation}

where $x_{i}$ is the frequency of the current $i$ time-window, $\mu_{k}$ and  $\sigma_{k}$ are mean and standard deviation of the previous $k$ time-windows respectively.

%\begin{itemize}
%	\item We calculate the number of messages of each time windows for each signal.
%	\item In our first approach, we compute max-min normalization and z-score normalization of all time windows for each signal. This approach is wrong because we do not know the next values in the streaming. We just know the previous values.
%	\item We compute the z-score normalization using past values for each signal independently. We do not use rolling/moving z-score because some signals have Null values of frequency. It is happen in smallest cities or states.
%\end{itemize}
\subsubsection{Inter-arrival Time}

To characterize the urgency of the messages during a time-window, we use our previous results \cite{DBLP:conf/fdia/Sarmiento17} to compute the \textit{Inter-arrival Time} which is defined as $d_{i} = t_{i+1} - t_{i}$, where $d_{i}$ denotes the difference between two consecutive social media messages $i$ and $i+1$ that arrived in moments $t_{i}$ and $t_{i+1}$ respectively. Using this definition and according to \cite{kalyanam2016prediction}, high-activity events have a high-frequency in the first bins represented by values $d_{i} \approx 0$.

In order to quantify a high-frequency in very small values of $d_{i}$, we compute the measures \textit{Skewness} and \textit{Kurtosis}, which represent the asymmetry and the tailedness of the shape of probability distribution respectively \cite{mardia1970measures}. Finally, we apply the equation \ref{eq:1} over \textit{Skewness} and \textit{Kurtosis} to calculate variation based on previous values.

%\begin{itemize}
%	\item Inter-arrival time represent the difference between 2 consecutive messages for one signal. In this case, a small difference between a set of messages represent the urgency or relevancy in a specific time (reference to plosone paper)
%	\item First, we compute inter-arrival time for each time windows independently. We summary results using quantiles, mean, min and max values. The problem with this approach is that we can not compare values between signals because each signal has different frequencies and behaviours along the time.
%	\item We compute skewness and kurtosis which represent a measure of the asymetry and "taildness" of the probability distribution respectively. In both case, we calculate the measures using previous values such as z-score normalization.
%\end{itemize}

%\subsection{Subsetting Signals}
%\begin{itemize}
%	\item Several signals have a null values along the time. For example, any city that is not a big city or capital city, does not a continuous frequency. To avoid this behaviour, we use cities just for compute the state frequency signal and later we discard them. Finally, we just monitor states and country hierarchy.
%	\item Furthermore, we remove gps signal because just 1-3 percent of messages have gps activated.
%\end{itemize}

\subsection{Geographic Spread}\label{sssec:geospread}

An emergency situation that affects and mobilizes response in a small area is defined as \textit{Focalized}, while a disaster with a large geographic impact is defined as \textit{Diffused} \cite{olteanu2015expect}. Using this definition, we extend this concept to represent neighborhoods between regions/states obtained from section \ref{sssec:geohie}. For that purpose, we create an \textit{Adjancency Matrix} $M$, where $M_{i,j} = 1$ represents if two states/regions are geographically connected and $M_{i,j} = 0$ if they are not connected. For instance, if an event is diffused (eg. earthquake), the detection should be in adjacent-locations independentely of metadata-level. In the other hand, if an event is focalized (eg. terrorist attack), just one-location should be detected but in different metadata-levels simultaneously.

\begin{table}
	\caption{List of earthquakes studied as ground truth, sorted by date.}
	\label{tab:eqs}
	\begin{tabular}{lccl}
		\toprule
		Country&Datetime (UTC)&Magnitude (Mw)& Language\\
		\midrule
		Italy & 2016-10-26 17:10:36 & 5.5 & Italian \\
		Italy & 2016-10-30 06:40:17 & 6.6 & Italian\\
		Chile & 2016-12-25 14:22:26 & 7.6 & Spanish\\
		Chile & 2017-04-23 02:36:06 & 5.9 & Spanish\\
		Chile & 2017-04-24 21:38:28 & 6.9 & Spanish\\
		\midrule
		%\bottomrule
	\end{tabular}
\end{table}

\section{Dataset}

In this section, data is collected from Twitter Public Streaming API, which allows access to subsets equal to $1\%$ of public status descriptions in real-time. With this tool, we can be retrieve either messages using a set of keywords or messages from specific locations setting a bounding box. In our approach, we get entire subsets of messages without use keywords or specific locations. Then, we retrieve random messages about any topic and any place in the world.

\subsection{Ground Truth}

We analyze five earthquakes with magnitudes between $5.5Mw$ and $7.6Mw$\footnote{Mw: the moment magnitude scale}, ocurring in Italian-speaking and Spanish-speaking countries between October 2016 and April 2017 (Table \ref{tab:eqs}). For that purpose, we collect $20$ millions of messages 12-hour before and after the emergency situation events.

According our proposal, we create both the gazetteer hierarchies\footnote{http://users.dcc.uchile.cl/$\sim$hsarmien/gazetteer.html} for each country and construct the signals based on each hierarchy and metadata-level. As a result of the number of messages of each signal (Table \ref{tab:message_signal}), we discard all signals related to city hierarchy since a great amount of small cities have zero frequency in a normal situation unlike to capital or metropolitan cities.

\begin{table}
	\caption{Number of messages by signal.}
	\label{tab:message_signal}
	\begin{tabular}{ccl}
		\toprule
		Hierarchy&Metadata-level&Messages\\
		\midrule
		All & All & 87,291 \\
		\midrule
	    \multirow{3}{*}{Country} & Tweet Text & 11,584 \\
	    & User Location & 25,313\\
	    & Tweet Text - User Location & 1,417\\	
		\bottomrule
		\multirow{3}{*}{State} & Tweet Text & 4,110 \\
		& User Location & 13,352\\
		& Tweet Text - User Location & 86\\
		\midrule
		\multirow{3}{*}{City} & Tweet Text & 1,415 \\
		& User Location & 8,971\\
		& Tweet Text - User Location & 20\\
		\bottomrule
	\end{tabular}
\end{table}

\subsubsection{Labeled Emergency Situations}

The exactly datetime event is obtained from National Seismology Agency in Chile\footnote{http://www.sismologia.cl/} and 
National Institute of Geophysics and Volcanology in Italy\footnote{http://www.ingv.it/it/}. For to label a time-window as \texttt{detection} class, we set detection label in those time-windows with positive variation in frequency, skewness and kurtosis respect to the normalization of the previous values (Figure \ref{fig:labeled}). Moreover, we include the three next time-windows after the event occurred to compensate the imbalance between classes.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{img/labeled2.png}
	\caption{Average variation in emergency situations between time-windows.}
	\label{fig:labeled}
\end{figure}

\section{Experiments}

Our filtering task can be seen as binary classification task. The positive class (\textit{detection label}) corresponds to messages related to instantaneous emergency situations, while the negative class (\textit{nothing label}) corresponds to the remaining or non-related to crisis situations.

In order to classifier messages, we employ traditional binary classifier Support Vector Machine \textit{(SVM)}. As a result of the analyzed data scattering (Figure \ref{fig:scatter}), we separate country and state in different datasets and set both kernels and classification parameters independentely. On the one hand, \textit{country} classifier uses a polynomial kernel and strict-parameters for gamma, cost and weights since that a great amount of messages are included in country hierarchy as an effect of the minor hierarchies. On the other hand, \textit{state/region} classifier uses a linear kernel with soft-weights and cost.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{img/scatter.png}
	\caption{Relationship between variables in country and state hierarchy. Red circles represent positive class (\textit{detection}) and blue circles represent negative class (\textit{nothing}).}
	\label{fig:scatter}
\end{figure}

Given that an emergency situation is not an usual event, we have an highly unbalanced data respect to the classes after labeled ($1\approx2\%$ of positive class corresponding to \textit{detection}). Therefore, we use \textit{under-sampling} \cite{lunardon2014rose} over country and state datasets increasing our possitive class to $15\approx18\%$. Additionally to validate our model, we use \textit{5-fold cross-validation} where one earthquake dataset is used as testing and the reamining earthquakes dataset as training.

Table \ref{tab:perfomance} shows the average results of our model applying 5-fold cross validation. In order provide a extended analysis about incorrect labels and time-windows, we include the metric \textit{False Positive Rate (FPR).}

\subsubsection{Independent Analysis of Hierarchies} 
Our first analysis is just considering the hierarchies as isolated detections. On the top of the Table \ref{tab:perfomance}, we show the results considering only the prediction over each label in our datasets. As noted above, the assignation from the lowest level (\textit{city}) to the highest (\textit{country}) in the gazetteer hierarchy generates high frequency of messages which cause multiples \textit{bursts} in our country signal for non-related emergency situations. This concept can explain the values of Precision (\textit{P}) and \textit{FPR}.

In addition to the analysis of number of detections by labels, we also study the number of detections by time-windows. For this analysis, we search the time-windows for each hierarchy where the all metadata-level are well classified with correct class. According to the results shown on the middle of the Table \ref{tab:perfomance}, the values of Precision, F1 and FPR have worst values than the analysis by label when we analyze country and state independentely. 
\begin{table}
	\caption{Average performance of 5-fold cross-validation by hierarchy and geographic spread (G.S.)}.
	\label{tab:perfomance}
	\centering
	\begin{tabular}{c|lcccc}
		\toprule
		&Hierarchy &P &R &F1 &FPR\\
		\midrule
		\parbox[t]{2mm}{\vspace{-0.2cm}\multirow{3}{*}{\rotatebox[origin=c]{90}{\small{label}}}} 
		& Country & 0.3 & 0.83 & 0.45& 0.14\\		
		& State &0.35& 0.83& 0.5 & 0.08\\
		\midrule
		\parbox[t]{2mm}{\vspace{0cm}\multirow{5}{*}{\rotatebox[origin=c]{90}{\small{time-window}}}}
		& Country & 0.15& 0.77 & 0.25 & 0.15\\
		& State & 0.17 & 0.88 & 0.29& 0.12\\
		& Country-State& 0.35 & 0.7 & 0.47 & 0.03\\
		& Country(2)-State with G.S. & 1 & 0.64 & 0.78 & 0 \\
		& Country(3)-State with G.S. & 1 & 0.47 & 0.64 & 0\\
		\bottomrule
	\end{tabular}
\end{table}

\subsubsection{Dependent Analysis of Hierarchies}
Our second analysis is considering the hierarchies as non-isolated detections. In the results explained above, we consider country and state hierarchy independently, which is not a correct analysis because an emergency situation affect to states and country at the same time. For this reason, we inspect the time-windows where all metadata-level for country and state hierarchy have a correct detection simultaneously. The results are shown in the row \textit{Country-State} in the Table \ref{tab:perfomance}. In contrast to the independent analysis of country and state, we improve the Precision, F1 and FPR values as a consequence of a smaller amount of non-related time-windows are assigned as detection. However, when we see the value obtained for FPR ($FPR = 0.03$), this rate represents an incorrect number of time-windows assigned as detection equal to $23$. This means that we have $22$ new emergency situations detected by our classifier.

\subsubsection{Geopraphic Spread Analysis}\label{sssec:geospreadanalysis}

Our third analysis is considering the hierarchies as non-isolated detections and using the Geographic Spread (G.S.). Using the \textit{Adjacency Matrix} to represent neighborhoods between regions/states, we consider as a correct detection those time-windows where the state/s classified as detection are defined as \textit{Focalized} or \textit{Diffused} and exist dependency between hierarchies. 
\begin{table*}
	\caption{Online evaluation by time-windows (T-W) using Country(2)-State with G.S. method}
	\label{tab:online1}
	\begin{tabular}{lclccp{3.5cm}l}
		\toprule
		Event &Detected T-W & T-W Before Event & T-W After Event & Delay(min) & Top 3 Bigrams\\
		\midrule
		Premier League Soccer Matches & 2& - & - & - & \small{(man, utd), (new, year), (happy, new)} \\
		Westminster Terrorist Attack& 13 & 0 &13 & 32& \small{(stay, safe), (terror, attack), (safe, everyone)}\\
		Manchester Terrorist Attack& 12& 1& 11& 23& \small{(ariana, grande), (incident, arena), (grande, concert)}\\
		London Terrorist Attack & 14 & 7 & 7 & 36 & \small{(stay, safe), (incident, bridge), (borough, market)}\\
		U.K. Elections& 5 & - & - & - & \small{(theresa, may), (vote, labour), (van, dijk)}\\
		Adele Live at Wembley& 9& 7&2&-& \small{(elland, road), (new, times), (phil, jackson)}\\
		England vs Eslovenia Soccer Match & 4& 4 & 0 & - & \small{(simon, brodkin), (join, us), (theresa, may)}\\
		Metallica Live at London& 4 & 4 & 0 & - & \small{(always, said), (chance, win), (carabao, cup)}\\
		\bottomrule
	\end{tabular}
\end{table*}

\begin{table*}
	\caption{Online evaluation by time-windows (T-W) using Country(3)-State with G.S. method}
	\label{tab:online2}
	\begin{tabular}{lclccp{3.5cm}l}
		\toprule
		Event &Detected T-W & T-W Before Event & T-W After Event & Delay(min) & Top 3 Bigrams\\
		\midrule
		Premier League Soccer Matches & 0& - & - & - & \hfill \break \\
		Westminster Terrorist Attack& 4 & 0 &4 & 32& \small{(terror, attack), (stay, safe), (terrorist, attack)}\\
		Manchester Terrorist Attack& 2& 0& 2& 23& \small{(ariana, grande), (praying, everyone), (everyone, affected)}\\
		London Terrorist Attack & 1 & 1 & 0 & - & \small{(ariana, grande), (around, world), (lady, gaga)} \\
		U.K. Elections& 0 & - & - & - & \hfill \break\\
		Adele Live at Wembley& 0& 0&0&-& \hfill \break\\
		England vs Eslovenia Soccer Match & 1& 1 & 0 & - & \small{(per, day), (menswear, sample), (closed, roads)}\\
		Metallica Live at London& 2 & 2 & 0 & - & \small{(happy, birthday), (chance, win), (always, said)}\\
		\bottomrule
	\end{tabular}
\end{table*}
In addition to the results of the dependency analysis explained above, we see that a large amount of time-windows ($\approx 82\%$) for country hierarchy have more than one metadata-level when exist a correct detection. This can be explained since an emergency situation produce a collective reaction on the level of body of the message (\textit{tweet text}), users sharing any messages with profile location in a specific country (\textit{user location}) or mixing both concepts (\textit{tweet text - user location}).

Considering the geographic spread by states and the number of metadata-levels by country hierarchy, we analyze the results shown on the bottom of the Table \ref{tab:perfomance}. On the one hand, the row with value equal to \textit{Country(2)-State with G.S.} represent the detection when we consider at least two metadata-levels for the country hierarchy and the geographic spread for states. In contrast to the the previous analyzes, we improve the values of the Precision, F1 and FPR. The last metrics is very important because there are not exist time-windows incorrectly assigned as emergency situations. Consequently, the Recall values decrease which means that our method remove some time-windows classified as detection. Beside the percent of emergency situations detected is equal to $100\%$ with a average delay equal to $10.4$ minutes ($min=6$, $max=14$) from the impact of the event to the detection.

On the other hand, \textit{Country(3)-State with G.S.} represent the detection when we consider three metadata-levels for country hierarchy and the geographic spread for states. Similar to  \textit{Country(2)-State with G.S.}, we improve the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with a average delay equal $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the detection.

\subsection{Online Evaluation}
For our evaluation in the Twitter Public Stream, we training our classifier with the five earthquakes identify in our ground truth. Futhermore, our evaluation dataset is formed by eight different events that ocurred in England between December 2016 and October 2017. For each event we consider the full-day in which they ocurred. The main goals of this evaluation is to know the capacity of our method to detect emergency situations and discard non-related to emergency situations that involve references of locations. Geographic spread analysis is used to evaluate our method because decrease the number of false positive detection. In the same way of the experiments in Section \ref{sssec:geospreadanalysis}, we compare the results using the two presented methods respect to the number of metadata-levels by country hierarchy.

As can be appreciated on Table \ref{tab:online1} and Table \ref{tab:online2}, we study three terrorist attacks and five high-impact real-world events related to soccer matches, music concerts and political elections. In the case of Premier League Soccer Matches and U.K Elections, we cannot identify the beginning of the event, since that in the first one there are many soccer matches during the analyzed day and in the second one there is not a specific start time. In order to know the topics when our method detect an event, we compute the Top 3 Bigrams in the detected time-windows. Also, we calculate the delay time just for emergency situations events.

On the one hand, the first evaluation \textit{Country(2)-State with G.S.} has full detection of the terrorist attacks with average delay time equal to $30.3$ minutes. These detections are related to the event given that the bigrams represent terms associated with crisis situations. However, the London Terrorist Attack has $50\%$ of the detected time-windows after the event, which means that there are seven time-windows non-related to emergency situations. Besides to the crisis situations analysis, we also study the number of detected time-windows in non-related to emergency situation events. In the same way, we have a large amount of misclassified time-windows that do not represent crisis situations as we can see in the Top 3 Bigrams for each non-related to event.

On the other hand, the second evaluation \textit{Country(3)-State with G.S.} decreases the number non-related to emergency situations detected as crisis situations. We can see three time-windows in two events detected as emergency situations (England vs Eslovenia, and Metallica Live at London). In these cases, the time-windows are detected before the event and corresponding a non-emergency situations according to the bigrams. Furthermore, when we analyzed the number of the detected emergency situations, two-third ($66\%$) of the events are detected correctly with average delay time equal to $30.3$ minutes . In the case of London Terrorist Attack, our method detects one time-window before the event but the bigrams describe that the detection do not correspond to crisis situations.

%Unlike to  \textit{Country(2)-State with G.S} method, \textit{Country(3)-State with G.S} method decreases the number of detection by time-windows given that latter is stricter than the first one respect to the number of country detection by metadata-level.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{img/delay.png}
	\caption{Delay time and number of locations in the first detection for diffused and focalized emergency situacions.}
	\label{fig:delay}
\end{figure}

\section{Discussion}

Our findings suggest that there is evidence to detect an emergency situation based on anomaly frequency of messages that contain locations for a specific country. Indeed, our method based on the number of metadata-levels by country hierarchy and geographic spread by state, detect $80\%$ of the events related to emergency situations as we could demostrate in our ground truth. Also, our method is independent of the textual features because we apply the model over different languages as spanish, italian and english. Furthermore, we test our model ind different type of events such as earthquakes and terrorist attacks
and also on different magnitudes (in the case of earthquakes) and affected people (e.g: Manchester Terrorist Attack vs Westminster Terrorist Attack).

However, when we apply our method in the online evaluation , we detect $66\%$ of the emergency situations that affected to England. This can explain because the signal are noise but many reason: the number of active users in United Kingdom\footnote{http://www.statista.com/statistics/242606/number-of-active-twitter-users-in-selected-countries/ visited on January 2018} which can affect the anomaly frequency of message since that exist a high daily average activity of the messages; similar locations in other countries (York $\approx$ New York); and soccer teams with names of cities (Manchester United, Liverpool). These issues also can affect the number of false positive detection in which in the case of England was $30\%$ of the non-related to emergency situations events.

Reggarding the geographic spread where we define an emergency situation as diffused or focalized, we find some evidences that differentiate them. In the case of diffused events, the delay time of the our first detection was less than $12$ minutes and in focalized events was greater than $30$ minutes (Figure \ref{fig:delay}). This can be explain given that, in diffused events such as earthquakes, a large amount of people are affected (thousand or millions) at the same time by an event which generate a collective reaction in social media in the locations where the event impact. In Figure \ref{fig:delay}, we can see that earthquakes have at least two detected locations in the first detection (except Italy EQ2). In contrast, focalized events have less amount of eyewitness (hundred or thousands) then when the users share messages in social media, the frequency not affect the average daily message of the country in the first minutes. This can be explain en Figure \ref{fig:delay} where the terrorist attacks have just 1 detected locations in the first detection.

Additionally, the delay time can be different according to many reason: datetime of the event (for example, during the early hours), little differences with the end of the current time-window, type of the affected locations (rural, urban cities) and the number of active users by locations.

%\begin{itemize}
%	\item explain that we dont need compute textual features as bigrams, top-k terms.
%	\item explain that we just need de gazetteer tree
%	\item explain when our methodology could be not working. eg. why not working correctly in england. country with many language. How event like soccer match are filtered and when not working
%	\item Delay time: explain difference between focalized and diffused event. People feel the event at the same time, activity in affected locations.
%	\item Number of locations affected in the first and second detection for focalized and diffused (plot)
%	\item Diffused: mercalli and locations detected.

%\end{itemize}

\section{Conclusion}

In this paper we have presented a methodology for detecting an emergency situation based on location for a specific country. This approach is independent of the textual features and can be used in different type of events and languages. We show that the users act self-organized in the affected locations like citizen sensors when an emergency situation occur. We furthermore have presented an analysis of geographic spread for different type of events, allowing characterize them. However, our experiment consider just a small portion of emergency situations which is not representive for all type of crisis situations according either the hazard type (natural or human-induced), temporal development (instantaneous or progressive) or geographic spread (diffused or focalized).

There are many things that can be improved our results. We will add Point of Interest to our gazetteer tree to increase the frequency by time-windows in each hierarchy. Furthermore, we will add more non-textual features as number of retweets and tweets, unique locations detected and special locations. We also plan study the relevance of the different metadata-levels and assign weights for each. Finally, we will create a web application to visualize in real-time events.

\begin{acks}
	The authors would like to thank Dr. Yuhua Li for providing the
	MATLAB code of the \textit{BEPS} method.

	The authors would also like to thank the anonymous referees for
	their valuable comments and helpful suggestions. The work is
	supported by the \grantsponsor{GS501100001809}{National Natural
		Science Foundation of
		China}{http://dx.doi.org/10.13039/501100001809} under Grant
	No.:~\grantnum{GS501100001809}{61273304}
	and~\grantnum[http://www.nnsf.cn/youngscientists]{GS501100001809}{Young
		Scientists' Support Program}.

\end{acks}

%\begin{itemize}
%	\item Add features for filtering detection: textual features. E.g: subjectivity, present tense, number of hashtag relative to the number of post, number of @s relative to the number of post, RT ratio, number of post that contain links.
%	\item Add features to classifier similar to above.
%	\item Evaluate the effect of gps signal.
%\end{itemize}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bio}

\end{document}

### 2018-01-26 paper.tex
Add reelated work

@@
 %\title{Detecting Emergency Situations Based on Locations in Twitter}
-\title{Self-Organized Citizen Sensors for Detecting Emergency Situations in Twitter}
+\title{Detecting Emergency Situations in Twitter Using Self-Organized Citizen Sensors}
 \titlenote{Produces the permission block, and
@@
 \affiliation{%
-  \institution{Deparment of Computer Science, University of Chile}
+  \institution{Department of Computer Science, University of Chile}
   \streetaddress{P.O. Box 1212}
@@
 	<concept>
-	<concept_id>10002951.10003227.10003236.10003239</concept_id>
-	<concept_desc>Information systems~Data streaming</concept_desc>
+	<concept_id>10002951.10003227.10003351.10003446</concept_id>
+	<concept_desc>Information systems~Data stream mining</concept_desc>
 	<concept_significance>500</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10002951.10003227.10003236</concept_id>
+	<concept_desc>Information systems~Spatial-temporal systems</concept_desc>
+	<concept_significance>300</concept_significance>
 	</concept>
@@
 
-\ccsdesc[500]{Information systems~Data streaming}
+\ccsdesc[500]{Information systems~Data stream mining}
+\ccsdesc[300]{Information systems~Spatial-temporal systems}
 
@@
 One of the main tasks during these situations is to detect a new real-event because as with most social media conversations, messages are often overcomed by irrelevant and redundant noise.
-In current works \cite{kumar2011tweettracker,caragea2011classifying,ashktorab2014tweedr,imranaidr2014,maldonado2017}, these tasks are solved with methods rely on keywords over \textit{Twitter Public Streaming API}\footnote{https://developer.twitter.com/en/docs/tweets/filter-realtime/overview}.
+In current works \cite{kumar2011tweettracker,caragea2011classifying,imranaidr2014,maldonado2017}, these tasks are solved with methods rely on keywords over \textit{Twitter Public Streaming API}\footnote{https://developer.twitter.com/en/docs/tweets/filter-realtime/overview}.
 The problem with these keyword-based methods is the need to train in specific domains for different type of events.
@@
 Additionally, several works related to crisis situations show a strong relationship between type of event (what) and the spatiotemporal dimensions (when and where).
-For example, we can see top trends during an earthquake related to locations \cite{mendoza2010twitter}, strong relationship between proximity to hurricane path and hurricane-related social media activity \cite{kryvasheyeu2016rapid}, extraction of locations and POI during floods \cite{lingad2013location} and mixing geographic information system (GIS) with geo-tagged message to improve disaster maping and real-time event tracking \cite{huang2015disastermapper}.
+For example, we can see top trends during an earthquake related to locations \cite{mendoza2010twitter}, strong relationship between proximity to hurricane path and hurricane-related social media activity \cite{kryvasheyeu2016rapid}, extraction of locations and POI during floods \cite{lingad2013location} and mixing geographic information system (GIS) with geo-tagged messages to improve disaster maping and real-time event tracking \cite{huang2015disastermapper}.
 
@@
 Next we present a complete description of our proposal divide in four parts.
-Thereafter, we provide a full description of our dataset and later we summarize our experimental validation and results over the ground truth and online evaluation.
+Thereafter, we provide a full description of our dataset and later we summarize our experimental validation over the ground truth and online evaluation.
 Finally, we deliver our discusions, conclusions and future work.
@@
 \section{Related Work}
-explain first works since 15 years ago
-\subsection{Emergency Situation}
-\subsubsection{Data Characterization}
-crisis lex, information nuggets, twitter under crisis, disaster related messages, practical extrac-
-tion of disaster-relevant, Identifying and Categorizing Disaster-Related Tweets,Microblogging
-During Two Natural Hazards Events: What Twitter May Contribute to Situational Aware-
-ness, Twitter Adoption and Use in Mass Convergence and Emergency Events, Twitter Under
-Crisis: Can we trust what we RT?
 
-\subsubsection{Crisis-Related Social Media Monitoring}
-tweedr, twicalli, ESA, aidr, twincident, EMERSE, twicalli, tweettracker, twittris
+Work related to this paper primarily arises from two areas: (1) crisis-related social media monitoring and (2) event detection based on locations.
 
-\subsection{Event Detection}
-burst detection (jheser), geo-spatial event detection, prediction and characterization of high-
-activity events,Emerging event detection in social networks with location sensitivity
+\subsection{Crisis-Related Social Media Monitoring}
 
-\subsection{Location Extraction}
-pinpointing, from the justin beiber’s heart, where in the world are you?
+Twitter has used extensively during emergency situations to extract and identify relevant information in recent years.
+However, social media communications during emergency situations are now so abundant that it is necessary to sift through millions of data point to find information that is most useful during a given event \cite{imran2015processing}.
+
+One of the main task related to emergency situations is to detect a new real-crisis event in social media.
+Most existing event-detection method described in the literature are based on keywords.
+For instance, \textit{TweetTracker} \cite{kumar2011tweettracker} presents a case study using tweets that discuss the Cholera outreak in Haiti.
+The primary mechanism for monitoring tweets is through specific keywords and hashtags filters related to the Haiti.
+To detect a new event, emerging trends are identified based on the analysis of older tweets.
+In the same way, \textit{EMERSE} \cite{caragea2011classifying} uses a set of keywords related to Haiti earthquake and applies random forests algorithm for classifyng messages.
+Likewise, \textit{Twicalli} system \cite{maldonado2017} introduces an unsupervised approach to detect earthquakes that only require a general list of keywords.
+In this work, detection is made monitoring \textit{zscore} variation through fixed time-windows.
+
+Researchers at CSIRO Australia propose \textit{ESA} \cite{cameron2012emergency,yin2012esa}, a system to detect disasters in Australia and New Zealand.
+This system are based on probabilistic method to identify bursty keywords and historical data to build a language model of word occurences.
+Alerts are identified if a term has a probability distribution that significantly deviates from the language model.
+Similarly to specific domains for a country, \textit{Twitcident} \cite{abel2012twitcident} system is developed for detecting incident which relies on emergency broadcasting service as police, fire departament or other public emergency services.
+\textit{Twitcident} framework translates the broadcasted message into an initial incident profile that is applied as query to collect messages from Twitter, where an incident profile is a set of weighted attribute-value pairs that describe the characteristics related to the incident.
+
+In order to describe different crisis-related terms to filter messages in Twitter, \textit{CrisisLex} is presented in \cite{olteanu2014}. In this work,
+authors collected six disasters which affected up to several million people.
+Data is collected from Twitter using two samples: a keyword-based sample and location-based sample.
+Based on datasets, they created a lexicon of the most frequent terms that appear in relevant messages posted during different types of crisis situations.
+
+\subsection{Event Detection Based on Locations}
+
+Additionally to crisis-related social media monitoring, event detection based on locations for other types of real-world events have been developed for which several unsupervised approaches have been proposed in the literature.
+\textit{Jasmine} \cite{watanabe2011jasmine} detects local events in the real-world using geolocation information from microblog documents.
+To detect such events, they identify a group of Twitter documents describing the same theme that were generated within a short time and a small geographic area.
+Similarly, \cite{unankard2015emerging} proposes an approach for the early detection of emerging hotspot events in social network with location sensivity.
+In this work, the authors identify strong correlations between user locations and event locations in detecting the emerging events using content similarity between clusters.
+In \cite{walther2013geo} real-world events are detected in a small scale with messages from the New York metropolitan area.
+There, clusters are created for each candidate event and evaluated using cluster score based on textual features (sentiment analysis, common theme, duplicate, etc) and other features (tweet count, unique coordinates, etc.).
+
+Supervised approaches have also been proposed.
+\textit{TEDAS} \cite{li2012tedas} detects, analyzes and identifies events using refined rules (e.g, keywords, hashtags) and classify messages based on content as well Twitter specific features as URLs, hashtags and mentions.
+Besides, locations information is extracted using both explicit geographical coordinates and implicit geographical references in the content.
+In the same way, \cite{becker2011beyond} proposes an online clustering technique, which continuosly clusters similar tweets and then classifies the clusters using Support Vector Machine algorithm.
+Finally, events (clusters) are classified into real-world events or nonevents.
+
+Summarizing, most previous works to detect an emergency situations rely on keywords using probabilistic temporal models for specific domains (eg.
+keywords or location).
+In general terms, this approach needs knowledge about the event and a new emerging topic referring to a crisis can not identify to detect an emergency situation.
+In the case of the works about event detection based on locations, they are designed to be applying in a small area and based on historical data.
+Regardless of which approach is implemented (supervised or unsupervised), textual features are the most used attribute to characterize an event and online clustering is the most common technique to create candidate events.
 
@@
 
-In order to provide a complete coverage of location-based detection of emergency situation, we proposed our approach which has four stages. The following information provides details of each stage.
+In order to provide a complete coverage of location-based detection of emergency situation, we propose an approach which has four stages as we can in Data Process module in Figure \ref{fig:diagram}.
+The following information provides details of each stage.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{img/diagram2.png}
+	\caption{Key components of the proposed approach.}
+	\label{fig:diagram}
+\end{figure}
 
@@
 Since that we consider users as citizen sensors, we filter messages depending of native language of each country using the attribute \textit{lang} retrieved from Tweet Metadata\footnote{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object}.
-For example, if we analyze England, we just consider messages in english.
+For example, if we analyze Italy, we just consider messages in italian.
 Finally, we remove user mentions, urls,  special characters and apply text tokenization .
-In this case, we do not remove hashtags or stopwords because some locations can be included in hashtags terms and some places can be contained stopword that can differentiate them from other similar non-related location terms.
+Also we do not remove hashtags or stopwords because some locations can be included in hashtags terms and some places can be contained stopword that can differentiate them from other similar non-related to locations terms.
 
@@
 Using the idea of \textit{Gazetteer as a Tree} presented in \cite{yin2014pinpointing} in which each place is associated with a canonical taxonomy node, we construct our gazetteer tree based on Geonames\footnote{http://download.geonames.org/export/dump/} and Wikipedia\footnote{http://www.wikipedia.org/} for each country to analyze.
-However, in \cite{yin2014pinpointing} the gazetteer hierarchy presents four-levels where the lowest level represents a specific POIs.
-In our approach, we use a subset of the gazetteer hierarchy with three-levels, in which the lowest level is represented by City since a large amount of users specify their location down to the city scale \cite{hecht2011tweets}.
+However, in \cite{yin2014pinpointing} the gazetteer hierarchy presents four levels where the lowest level represents a specific POIs.
+In our approach, we use a subset of the gazetteer hierarchy with three levels, in which the lowest level is represented by City since a large amount of users specify their location down to the city scale \cite{hecht2011tweets}.
 For example, if we have the \textit{city:Manchester}, we associate this location with \textit{region-state:North West} and also with \textit{country:England}.
@@
 The structure of the Tweet Metadata allows to obtain information about the message and the user who shares a message.
-Considering the geographical hierarchy explained in the previous section, we search these locations on different levels of metadata and create one-signal for each as following:
+Considering the geographical hierarchy explained in the previous section, we search these locations on different levels of metadata and create one signal for each as following:
 
@@
 	\item Tweet Text - User Location: location is mentioned in the attribute \texttt{text} of tweet object and also location is mentioned in the attribute \texttt{location} inside the \texttt{user object}.
-This means that the location is mentioned on the body on the message and the user who share message has the same location in his profile.
+This means that the location is mentioned on the body of the message and the user who share message has the same location in his profile.
 In this case, tweet text and user location can be different in the smallest hierarchy but in the highest level can be equal.
@@
 
-In this way, mixing geographical hierarchy and locations in micro-blog metadata, we create $NxM$ signals where $N$ is the number of locations obtained by gazetteer tree and $M$ is the number of metadata-level extracted from Tweet Object.
+In this way, mixing geographical hierarchy and locations in micro-blog metadata, we create $NxM$ signals where $N$ is the number of locations obtained by gazetteer tree and $M$ is the number of metadata-levels extracted from Tweet Object.
 In instance, we create a signal for \textit{city:Manchester} and we find this hierarchy in \textit{metadata:Tweet Text} and also in \textit{metadata:User Location}.
@@
 \subsection{Time-Windows}
-In this stage we address the problems of how to divide and determine the time-windows size to detect a new emergency situation and what variables by time-window allow it.
+In this stage we address the problems of how to divide and determine the time-windows size to detect a new emergency situation and what features by time-window allow it.
 
-\subsubsection{Determining optimal window size}
+\subsubsection{Determining Optimal Window Size}
 
 According to \cite{guzman2013line}: ``If the window size is to small, the ocurrence of empty windows for a term increases, making the noise rate increate and frequency rate tend to zero.
-On the other hand, if the window size is too large, the stability of the signal becomes constant and busrty keyword detection is delayed''.
+On the other hand, if the window size is too large, the stability of the signal becomes constant and bursty keyword detection is delayed''.
 Using this definition, we divide our signals into windows of six minutes because it divides a 24-hour day exactly, making the analysis easier to understand and to compare from different days.
@@
 An emergency situation that affects and mobilizes response in a small area is defined as \textit{Focalized}, while a disaster with a large geographic impact is defined as \textit{Diffused} \cite{olteanu2015expect}.
-Using this definition, we extend this concept to represent neighborhoods between regions/states obtained from section \ref{sssec:geohie}.
-For that purpose, we create an \textit{Adjancency Matrix} $M$, where $M_{i,j} = 1$ represents if two states/regions are geographically connected and $M_{i,j} = 0$ if they are not connected.
+Using this definition, we extend this concept to represent neighborhoods between locations obtained from section \ref{sssec:geohie}.
+For that purpose, we create an \textit{Adjancency Matrix} $M$, where $M_{i,j} = 1$ represents if two locations are geographically connected and $M_{i,j} = 0$ if they are not connected.
 For instance, if an event is diffused (eg.
@@
 In the other hand, if an event is focalized (eg.
-terrorist attack), just one-location should be detected but in different metadata-levels simultaneously.
+terrorist attack), just one location should be detected but in different metadata-levels simultaneously.
 
@@
 
-In this section, data is collected from Twitter Public Streaming API, which allows access to subsets equal to $1\%$ of public status descriptions in real-time.
-With this tool, we can be retrieve either messages using a set of keywords or messages from specific locations setting a bounding box.
+Data is collected from Twitter Public Streaming API, which allows access to subsets equal to $1\%$ of public status descriptions in real-time.
+With this tool, we can retrieve either messages using a set of keywords or messages from specific locations setting a bounding box.
 In our approach, we get entire subsets of messages without use keywords or specific locations.
@@
 National Institute of Geophysics and Volcanology in Italy\footnote{http://www.ingv.it/it/}.
-For to label a time-window as \texttt{detection} class, we set detection label in those time-windows with positive variation in frequency, skewness and kurtosis respect to the normalization of the previous values (Figure \ref{fig:labeled}).
-Moreover, we include the three next time-windows after the event occurred to compensate the imbalance between classes.
+With the purpose of to labeling a time-window as positive class (detection), we set as detection those time-windows with positive variation in frequency, skewness and kurtosis respect to the normalization of the previous values.
+Moreover and according to (Figure \ref{fig:labeled}), we include the three next time-windows after the event to compensate the imbalance between classes given that later to these number of time-windows, the variation in the features decrease.
 
@@
 	\includegraphics[width=\columnwidth]{img/scatter.png}
-	\caption{Relationship between variables in country and state hierarchy. Red circles represent positive class (\textit{detection}) and blue circles represent negative class (\textit{nothing}).}
+	\caption{Relationship between features in country and state hierarchy. Red circles represent positive class (\textit{detection}) and blue circles represent negative class (\textit{nothing}).}
 	\label{fig:scatter}
@@
 On the top of the Table \ref{tab:perfomance}, we show the results considering only the prediction over each label in our datasets.
-As noted above, the assignation from the lowest level (\textit{city}) to the highest (\textit{country}) in the gazetteer hierarchy generates high frequency of messages which cause multiples \textit{bursts} in our country signal for non-related emergency situations.
+As noted above, the assignation from the lowest level (\textit{city}) to the highest (\textit{country}) in the gazetteer hierarchy generates high frequency of messages which cause multiples \textit{bursts} in our country signal for non  emergency situations.
 This concept can explain the values of Precision (\textit{P}) and \textit{FPR}.
@@
 In addition to the analysis of number of detections by labels, we also study the number of detections by time-windows.
-For this analysis, we search the time-windows for each hierarchy where the all metadata-level are well classified with correct class.
-According to the results shown on the middle of the Table \ref{tab:perfomance}, the values of Precision, F1 and FPR have worst values than the analysis by label when we analyze country and state independentely.
+For this analysis, we search the time-windows for each hierarchy where the all metadata-levels are well classified with correct class.
+According to the results shown on the middle of the Table \ref{tab:perfomance}, when we analyze country and state independentely the values of Precision, F1 and FPR have worst values than the analysis by label .
 \begin{table}
@@
 The results are shown in the row \textit{Country-State} in the Table \ref{tab:perfomance}.
-In contrast to the independent analysis of country and state, we improve the Precision, F1 and FPR values as a consequence of a smaller amount of non-related time-windows are assigned as detection.
+In contrast to the independent analysis of country and state, we improve the Precision, F1 and FPR values as a consequence of a smaller amount of the time-windows related to non emergency situations are assigned as detection.
 However, when we see the value obtained for FPR ($FPR = 0.03$), this rate represents an incorrect number of time-windows assigned as detection equal to $23$.
-This means that we have $22$ new emergency situations detected by our classifier.
+This means that we have $23$ new emergency situations detected by our classifier.
 
@@
 
-Our third analysis is considering the hierarchies as non-isolated detections and using the Geographic Spread (G.S.).
+Our third analysis is considering the hierarchies as non-isolated detections and applying the Geographic Spread (G.S.).
 Using the \textit{Adjacency Matrix} to represent neighborhoods between regions/states, we consider as a correct detection those time-windows where the state/s classified as detection are defined as \textit{Focalized} or \textit{Diffused} and exist dependency between hierarchies.
+
 \begin{table*}
@@
 	\label{tab:online1}
-	\begin{tabular}{lclccp{3.5cm}l}
+	\begin{tabular}{lccccp{3.5cm}c}
 		\toprule
-		Event &Detected T-W & T-W Before Event & T-W After Event & Delay(min) & Top 3 Bigrams\\
+		Event &Detected T-W & T-W Before Event & T-W After Event & Delay (min) & Top 3 Bigrams\\
 		\midrule
@@
 	\label{tab:online2}
-	\begin{tabular}{lclccp{3.5cm}l}
+	\begin{tabular}{lccccp{3.5cm}c}
 		\toprule
-		Event &Detected T-W & T-W Before Event & T-W After Event & Delay(min) & Top 3 Bigrams\\
+		Event &Detected T-W & T-W Before Event & T-W After Event & Delay (min) & Top 3 Bigrams\\
 		\midrule
@@
 \end{table*}
-In addition to the results of the dependency analysis explained above, we see that a large amount of time-windows ($\approx 82\%$) for country hierarchy have more than one metadata-level when exist a correct detection.
+In addition to the results of the dependency analysis explained above, we see that a large amount of time-windows for country hierarchy ($\approx 82\%$) have more than one metadata-level when exist a correct detection.
 This can be explained since an emergency situation produce a collective reaction on the level of body of the message (\textit{tweet text}), users sharing any messages with profile location in a specific country (\textit{user location}) or mixing both concepts (\textit{tweet text - user location}).
@@
 In contrast to the the previous analyzes, we improve the values of the Precision, F1 and FPR.
-The last metrics is very important because there are not exist time-windows incorrectly assigned as emergency situations.
+The last metrics is very important because there are no time-windows incorrectly assigned as emergency situations.
 Consequently, the Recall values decrease which means that our method remove some time-windows classified as detection.
-Beside the percent of emergency situations detected is equal to $100\%$ with a average delay equal to $10.4$ minutes ($min=6$, $max=14$) from the impact of the event to the detection.
+Beside the percent of emergency situations detected is equal to $100\%$ with a average delay equal to $10.4$ minutes ($min=6$, $max=14$) from the impact of the event to the first detection.
 
 On the other hand, \textit{Country(3)-State with G.S.} represent the detection when we consider three metadata-levels for country hierarchy and the geographic spread for states.
-Similar to  \textit{Country(2)-State with G.S.}, we improve the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with a average delay equal $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the detection.
+Similar to  \textit{Country(2)-State with G.S.}, we improve the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with a average delay equal to $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the first detection.
 
 \subsection{Online Evaluation}
-For our evaluation in the Twitter Public Stream, we training our classifier with the five earthquakes identify in our ground truth.
+For our evaluation in the Twitter Public Stream, we training a classifier with five earthquakes identified in our ground truth.
 Futhermore, our evaluation dataset is formed by eight different events that ocurred in England between December 2016 and October 2017.
 For each event we consider the full-day in which they ocurred.
-The main goals of this evaluation is to know the capacity of our method to detect emergency situations and discard non-related to emergency situations that involve references of locations.
-Geographic spread analysis is used to evaluate our method because decrease the number of false positive detection.
+The main goals of this evaluation is to know the capacity of our method to detect emergency situations and discard non-related to emergency situations events that involve references of locations.
+Geographic spread analysis is used to evaluate our method because decrease the number of false positives detection.
 In the same way of the experiments in Section \ref{sssec:geospreadanalysis}, we compare the results using the two presented methods respect to the number of metadata-levels by country hierarchy.
@@
 As can be appreciated on Table \ref{tab:online1} and Table \ref{tab:online2}, we study three terrorist attacks and five high-impact real-world events related to soccer matches, music concerts and political elections.
-In the case of Premier League Soccer Matches and U.K Elections, we cannot identify the beginning of the event, since that in the first one there are many soccer matches during the analyzed day and in the second one there is not a specific start time.
+In the case of Premier League Soccer Matches and U.K Elections, we can not identify the beginning of the event, since that in the first one there are many soccer matches during the analyzed day and in the second one there is no a specific start time.
 In order to know the topics when our method detect an event, we compute the Top 3 Bigrams in the detected time-windows.
@@
 
-On the other hand, the second evaluation \textit{Country(3)-State with G.S.} decreases the number non-related to emergency situations detected as crisis situations.
+On the other hand, the second evaluation \textit{Country(3)-State with G.S.} decreases the number non-related to emergency situations events detected as crisis situations.
 We can see three time-windows in two events detected as emergency situations (England vs Eslovenia, and Metallica Live at London).
-In these cases, the time-windows are detected before the event and corresponding a non-emergency situations according to the bigrams.
+In these cases, the time-windows are detected before the event and corresponding a non emergency situations according to the bigrams.
 Furthermore, when we analyzed the number of the detected emergency situations, two-third ($66\%$) of the events are detected correctly with average delay time equal to $30.3$ minutes .
-In the case of London Terrorist Attack, our method detects one time-window before the event but the bigrams describe that the detection do not correspond to crisis situations.
+In the case of London Terrorist Attack, our method detects one time-window before the event but the bigrams describe that the detection do not correspond to crisis situation.
 
@@
 Also, our method is independent of the textual features because we apply the model over different languages as spanish, italian and english.
-Furthermore, we test our model ind different type of events such as earthquakes and terrorist attacks
-and also on different magnitudes (in the case of earthquakes) and affected people (e.g: Manchester Terrorist Attack vs Westminster Terrorist Attack).
+Furthermore, we test our model in different types of events such as earthquakes and terrorist attacks, and also on different magnitudes (in the case of earthquakes) and number of affected people (e.g: Manchester Terrorist Attack vs Westminster Terrorist Attack).
 
@@
 In the case of diffused events, the delay time of the our first detection was less than $12$ minutes and in focalized events was greater than $30$ minutes (Figure \ref{fig:delay}).
-This can be explain given that, in diffused events such as earthquakes, a large amount of people are affected (thousand or millions) at the same time by an event which generate a collective reaction in social media in the locations where the event impact.
+This can be explain given that, in diffused events such as earthquakes, a large amount of people are affected (thousand or millions) at the same time by an event which generate a collective reaction in social media in the locations where the event impacted.
 In Figure \ref{fig:delay}, we can see that earthquakes have at least two detected locations in the first detection (except Italy EQ2).
@@
 In this paper we have presented a methodology for detecting an emergency situation based on location for a specific country.
-This approach is independent of the textual features and can be used in different type of events and languages.
+This approach is independent of the textual features and can be used in different types of events and languages.
 We show that the users act self-organized in the affected locations like citizen sensors when an emergency situation occur.
-We furthermore have presented an analysis of geographic spread for different type of events, allowing characterize them.
+We furthermore have presented an analysis of geographic spread for different types of events, allowing characterize them.
 However, our experiment consider just a small portion of emergency situations which is not representive for all type of crisis situations according either the hazard type (natural or human-induced), temporal development (instantaneous or progressive) or geographic spread (diffused or focalized).
@@
 \begin{acks}
-	The authors would like to thank Dr. Yuhua Li for providing the
-	MATLAB code of the \textit{BEPS} method.
-
-	The authors would also like to thank the anonymous referees for
-	their valuable comments and helpful suggestions. The work is
-	supported by the \grantsponsor{GS501100001809}{National Natural
-		Science Foundation of
-		China}{http://dx.doi.org/10.13039/501100001809} under Grant
-	No.:~\grantnum{GS501100001809}{61273304}
-	and~\grantnum[http://www.nnsf.cn/youngscientists]{GS501100001809}{Young
-		Scientists' Support Program}.
 
 \end{acks}
-
-%\begin{itemize}
-%	\item Add features for filtering detection: textual features.
-E.g: subjectivity, present tense, number of hashtag relative to the number of post, number of @s relative to the number of post, RT ratio, number of post that contain links.
-%	\item Add features to classifier similar to above.
-%	\item Evaluate the effect of gps signal.
-%\end{itemize}
 

### 2018-02-05 paper.tex
add barbara's modifications

@@
 %\setcopyright{cagovmixed}
+%\usepackage{setspace}
+\pagestyle{empty} 
+%\doublespacing
 
@@
 %Conference
-\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El Paso, Texas USA} 
-\acmYear{1997}
+%\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El Paso, Texas USA} 
+%\acmYear{1997}
 %\copyrightyear{2016}
@@
 %\title{Detecting Emergency Situations Based on Locations in Twitter}
-\title{Detecting Emergency Situations in Twitter Using Self-Organized Citizen Sensors}
-\titlenote{Produces the permission block, and
-  copyright information}
+%\title{Detecting Emergency Situations in Twitter Using Self-Organized Citizen Sensors}
+\title{Domain-Independent Detection of Emergency Situations\\Based on Social Activity Related to Geolocations}
+
+%\titlenote{Produces the permission block, and
+ % copyright information}
+
 %\subtitle{Extended Abstract}
@@
   \institution{Department of Geophysics, University of Chile}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
+%  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
   \city{Santiago} 
@@
 Our method use a SVM classifier that is independent of the text language of the messages and can be training and testing among different languages and domains.
-
-Most methods to detect emergency situations using Twitter rely on keyword.
-The problem of keyword-based methods is the need to train models in specific domains, in multiple languages, and for different types of events, for example: earthquakes, terrorist attacks, floods, etc.
-In contrast, our proposal is to rather use the recurring references of a location in the metadata of microblogging messages, where users act self-organized citizen sensors in real-locations.
-In order to identify an emergency situation, we characterize  such events  through frequencies, and probability distributions of the interarrival time of the messages.
-Our method uses a SVM classifier which is independent of the language of the text, and textual features.
-Finally, we describe an emergency situation according to geographic impact based on the propagation of the messages in social media between neighboring locations.
+%
+Most methods for detecting emergency situations using Twitter rely on identifying features within messages that contain domain-specific keywords.
+However, keyword-based methods require models to be trained on historical data of specific domains, in multiple languages, and for different types of events (e.g., earthquakes, floods, wildfires, etc.).
+%
+In addition to being costly, these approaches may fail to detect previously unexpected situations, such as uncommon catastrophes or terrorist attacks.
+%
+Nevertheless, collective mentions of keywords are not the only type of self-organizing phenomena that may arise when a real-world extreme situation occurs.
+%
+Just as physical sensors become activated when stimulated, localized citizen sensors will also react in a similar manner.  
+%
+To leverage this information, we propose to use self-organized geolocation related activity to identify emergency situations.
+%
+We propose to detect such events by tracking the frequencies, and probability distributions of the interarrival time of the messages related to specific locations.
+%
+Using an off-the-shelf classifier that is independent of domain-specific features, we study and describe emergency situations based solely on location-based features in messages.
+%
+Our findings indicate that anomalies in location-related social media user activity indeed provide information for automatically detecting emergency situations independent of their domain.
 \end{abstract}
@@
 \section{Introduction}
-Social media has become a major channel for communication during high-impact real-world events, for example: elections, sports events, emergency situations, etc.
-In any case, users act as ``citizen sensors'' where they share and post their mood, opinions, photos, videos and geographical point of interest (POI).
+Social media has become a major communication channel during high-impact real-world events, such as elections, sports events, emergency situations, among others.
+In each case, users act as ``citizen sensors'' sharing and posting their mood, opinion, photos, videos and geographical points of interest.
 
-During emergency situations, traditional media suffer structural damage and real-time communications could be disrupted.
-Instead, microblogging has played a critical role in the last fifteen years since it has allowed users share real-time information from people local to the incident as current status, casualties, damages and alerts \cite{kumar2011tweettracker,imran2013extracting,stowe2016identifying,reuterfifteen}.
-For this reason, researchers have studied the behaviour during these events for to detect, summarize and classifyng messages with the goal to help authorities and the general public with situational awareness to act fast and conscientiously during a crisis situations.
+During emergency situations, traditional media may suffer infrastructure issues and real-time communications could be disrupted.
+Instead, microblogging has played a critical role over the last fifteen years allowing users to share real-time information from people local to the incident, such as status updates, casualties, damages and alerts \cite{kumar2011tweettracker,imran2013extracting,stowe2016identifying,reuterfifteen}.
+For this reason, researchers have studied user behavior during these events to detect, summarise and classify messages with the goal of helping authorities, and the general public, with situational awareness to provide fast and conscientious responses during crisis situations.
 
-Twitter, a microblogging which allows to share short messages (called \textit{tweets}) and is currently used world-wide by over $300$ million users\footnote{http://www.statista.com/statistics/282087/number-of-monthly-active-twitter-users/}.
-About $80\%$ of Twitter users access from mobile services which contribute to the inmediacy of diffusion of information specially during a crisis situations \cite{castillo2016big}.
-One of the main tasks during these situations is to detect a new real-event because as with most social media conversations, messages are often overcomed by irrelevant and redundant noise.
-In current works \cite{kumar2011tweettracker,caragea2011classifying,imranaidr2014,maldonado2017}, these tasks are solved with methods rely on keywords over \textit{Twitter Public Streaming API}\footnote{https://developer.twitter.com/en/docs/tweets/filter-realtime/overview}.
-The problem with these keyword-based methods is the need to train in specific domains for different type of events.
-\cite{olteanu2014} generate set of keywords with different datasets but sometimes arise specific terms for one event, for example \textit{\#eqnz} for Earthquake in New Zealand or \textit{\#pabloph} for Typhoon Pablo in Philippines \cite{potts2011tweeting,bruns2012local,karimi2013classifying}.
-Furthermore, these set of keywords are trained for a unique language and can not be used in others.
+Twitter is a microblogging platform that allows users to share short messages (called \textit{tweets}) and is currently used worldwide by over $300$ million people.\footnote{http://www.statista.com/statistics/282087/number-of-monthly-active-twitter-users/} About $80\%$ of Twitter users access from mobile devices, which contributes to the immediacy of diffusion of information, especially during crisis situations \cite{castillo2016big}.
+One of the most important tasks during these situations, is to timely detect incoming real-world events.
+This is because, as with most social media conversations, messages are often overridden with irrelevant and redundant noise.
+In current works \cite{kumar2011tweettracker,caragea2011classifying,imranaidr2014,maldonado2017}, these tasks are solved with methods that rely on keyword based filters on the \textit{Twitter Public Streaming API}.\footnote{https://developer.twitter.com/en/docs/tweets/filter-realtime/overview} The problem with keyword-based methods is the need to train in specific domains for different type of events.
+For instance, \citeauthor{olteanu2014} \cite{olteanu2014} generate a set of keywords based on different datasets.
+However, this is not sufficient for cases in which specific and previously unseen terms arise for a particular event (e.g., \textit{\#eqnz} for Earthquakes in New Zealand, or \textit{\#pabloph} for Typhoon Pablo in Philippines) \cite{potts2011tweeting,bruns2012local,karimi2013classifying}.
+Furthermore, these sets of keywords are commonly obtained a particular language and will not work in others.
 
-Additionally, several works related to crisis situations show a strong relationship between type of event (what) and the spatiotemporal dimensions (when and where).
-For example, we can see top trends during an earthquake related to locations \cite{mendoza2010twitter}, strong relationship between proximity to hurricane path and hurricane-related social media activity \cite{kryvasheyeu2016rapid}, extraction of locations and POI during floods \cite{lingad2013location} and mixing geographic information system (GIS) with geo-tagged messages to improve disaster maping and real-time event tracking \cite{huang2015disastermapper}.
+Additionally, several works related to crisis situations show a strong relationship between a type of event (what) and the spatiotemporal dimensions (when and where).
+For example, it has been observed that top trends during earthquakes are related to location mentions \cite{mendoza2010twitter}, also, that there is a strong relationship between the proximity to a hurricane path and hurricane-related social media activity \cite{kryvasheyeu2016rapid}.
+Other works have addressed the extraction of locations and points of interest during floods \cite{lingad2013location}, as well as mixing geographic information system (GIS) information with geo-tagged messages to improve disaster mapping and real-time event tracking \cite{huang2015disastermapper}.
 
-Based on previous works, the question we address here is: \textit{Is there evidence to detect an emergency situation based on anomaly frequency of messages that contain locations?}.
+Based on previous works, the question we address here is: \textit{Is there evidence to detect an emergency situation based on anomalies in messages related to contain locations?}.
 
-Here, we propose a method based on recurring references of country locations in message's metadata.
-For this task, we create a gazetteer tree based on the hierarchy of the specific country, divide the messages into fixed time-windows and compute the frequency, and the probability distributions of the inter-arrival time of the messages for each geographical hierarchy.
+Here, we propose a method based on recurring references to country level locations in message metadata.
+For this task, we create a gazetteer tree based on the hierarchy of the specific country, divide the messages into fixed time-windows and compute the frequency and the probability distributions of the interarrival time of the messages for each geographical hierarchy.
 To detect an emergency situation, we train a SVM classifier by each hierarchy and apply a geographic spread to filter false positives detection considering that an emergency situation can be \textit{Focalized} or \textit{Diffused}.
 
-Our main contribution is to create a methodology that detecting instantaneous emergency situations just using the locations for a specific country.
-Furthermore, our classifier do not depend of the language to detect a new event since that do not use textual features as input.
-Moreover, we characterize crisis situations that affect a small or large geographic area.
+Our main contribution is to create a methodology that detects instantaneous emergency situations just by using the location related information for a specific country.
+Furthermore, our classifiers do not depend on language to detect a new event since they do not use textual features as input.
+Moreover, we characterize crisis situations that affect small or large geographical areas.
 
@@
 Next we present a complete description of our proposal divide in four parts.
-Thereafter, we provide a full description of our dataset and later we summarize our experimental validation over the ground truth and online evaluation.
-Finally, we deliver our discusions, conclusions and future work.
+Thereafter, we provide a full description of our dataset and later we summarise our experimental validation over the ground truth and on-line evaluation.
+Finally, we deliver our discussions, conclusions and future work.
 
@@
 
-Work related to this paper primarily arises from two areas: (1) crisis-related social media monitoring and (2) event detection based on locations.
+We discuss relevant work for our proposal, primary from two areas: (1) crisis-related social media monitoring and (2) event detection based on locations.
 
@@
 
-Twitter has used extensively during emergency situations to extract and identify relevant information in recent years.
-However, social media communications during emergency situations are now so abundant that it is necessary to sift through millions of data point to find information that is most useful during a given event \cite{imran2015processing}.
+Twitter has been used extensively during emergency situations to extract and identify relevant information.
+However, social media communications during emergency situations have become so abundant that it is necessary to sift through millions of data points to find information that is most useful during a given event \cite{imran2015processing}.
 
-One of the main task related to emergency situations is to detect a new real-crisis event in social media.
-Most existing event-detection method described in the literature are based on keywords.
-For instance, \textit{TweetTracker} \cite{kumar2011tweettracker} presents a case study using tweets that discuss the Cholera outreak in Haiti.
+One of the main tasks related to emergency situations is to detect a new real-crisis event in social media.
+Most existing event-detection methods described in the literature are based on keywords.
+For instance, \textit{TweetTracker} \cite{kumar2011tweettracker} presented a case study using tweets that discuss a cholera outbreak in Haiti.
 The primary mechanism for monitoring tweets is through specific keywords and hashtags filters related to the Haiti.
 To detect a new event, emerging trends are identified based on the analysis of older tweets.
-In the same way, \textit{EMERSE} \cite{caragea2011classifying} uses a set of keywords related to Haiti earthquake and applies random forests algorithm for classifyng messages.
-Likewise, \textit{Twicalli} system \cite{maldonado2017} introduces an unsupervised approach to detect earthquakes that only require a general list of keywords.
-In this work, detection is made monitoring \textit{zscore} variation through fixed time-windows.
+In the same way, \textit{EMERSE} \cite{caragea2011classifying} used a set of keywords related to the Haiti earthquake and applies a random forests algorithm for classifying messages.
+Likewise, the \textit{Twicalli} system \cite{maldonado2017} introduced an unsupervised approach to detect earthquakes that only requires a general list of keywords.
 
-Researchers at CSIRO Australia propose \textit{ESA} \cite{cameron2012emergency,yin2012esa}, a system to detect disasters in Australia and New Zealand.
-This system are based on probabilistic method to identify bursty keywords and historical data to build a language model of word occurences.
+Researchers at CSIRO Australia proposed \textit{ESA} \cite{cameron2012emergency,yin2012esa}, a system to detect disasters in Australia and New Zealand.
+This system is based on a probabilistic method to identify bursty keywords and historical data to build a language model of word occurrences.
 Alerts are identified if a term has a probability distribution that significantly deviates from the language model.
-Similarly to specific domains for a country, \textit{Twitcident} \cite{abel2012twitcident} system is developed for detecting incident which relies on emergency broadcasting service as police, fire departament or other public emergency services.
-\textit{Twitcident} framework translates the broadcasted message into an initial incident profile that is applied as query to collect messages from Twitter, where an incident profile is a set of weighted attribute-value pairs that describe the characteristics related to the incident.
+Similarly, \textit{Twitcident} \cite{abel2012twitcident} system was developed for detecting incidents which relies on emergency broadcasting services, such as the police, the fire department and other public emergency services.
+The \textit{Twitcident} framework translates the broadcasted message into an initial incident profile that is applied as query to collect messages from Twitter, where an incident profile is a set of weighted attribute-value pairs that describe the characteristics related to the incident.
 
-In order to describe different crisis-related terms to filter messages in Twitter, \textit{CrisisLex} is presented in \cite{olteanu2014}. In this work,
-authors collected six disasters which affected up to several million people.
-Data is collected from Twitter using two samples: a keyword-based sample and location-based sample.
+\textit{CrisisLex} \cite{olteanu2014} is introduced to extract different crisis-related terms to filter messages in Twitter. In this work,
+authors collected six disasters which affected several millions of people.
+Data was collected from Twitter using two samples: a keyword-based sample and location-based sample.
 Based on datasets, they created a lexicon of the most frequent terms that appear in relevant messages posted during different types of crisis situations.
+
+Our current work, on the other hand, is based on detecting events by monitoring \textit{zscore} variations through fixed time-windows.
 
@@
 
-Additionally to crisis-related social media monitoring, event detection based on locations for other types of real-world events have been developed for which several unsupervised approaches have been proposed in the literature.
-\textit{Jasmine} \cite{watanabe2011jasmine} detects local events in the real-world using geolocation information from microblog documents.
-To detect such events, they identify a group of Twitter documents describing the same theme that were generated within a short time and a small geographic area.
-Similarly, \cite{unankard2015emerging} proposes an approach for the early detection of emerging hotspot events in social network with location sensivity.
-In this work, the authors identify strong correlations between user locations and event locations in detecting the emerging events using content similarity between clusters.
-In \cite{walther2013geo} real-world events are detected in a small scale with messages from the New York metropolitan area.
-There, clusters are created for each candidate event and evaluated using cluster score based on textual features (sentiment analysis, common theme, duplicate, etc) and other features (tweet count, unique coordinates, etc.).
+In addition to crisis-related social media monitoring, there are also some unsupervised event detection approaches based on location information.
+\textit{Jasmine} \cite{watanabe2011jasmine} detected local real-world events using geolocation information from microblog documents.
+To detect such events, they identify a group of tweets that describe a particular theme, which are generated within a short time frame and a same geographic area.
+Similarly, \citeauthor*{unankard2015emerging} \cite{unankard2015emerging} proposed an approach for early detection of emerging hotspot events in social networks with location sensivity.
+In this work, the authors identified strong correlations between user locations and event locations when detecting emerging events using content similarity between clusters.
+In the work of \citeauthor{walther2013geo} \cite{walther2013geo} real-world events were detected in a small scale with messages from the New York metropolitan area.
+There, clusters were created for each candidate event and evaluated using cluster score based on textual features (sentiment analysis, common theme, duplicate, etc) and other features (tweet count, unique coordinates, etc.).
 
 Supervised approaches have also been proposed.
-\textit{TEDAS} \cite{li2012tedas} detects, analyzes and identifies events using refined rules (e.g, keywords, hashtags) and classify messages based on content as well Twitter specific features as URLs, hashtags and mentions.
-Besides, locations information is extracted using both explicit geographical coordinates and implicit geographical references in the content.
-In the same way, \cite{becker2011beyond} proposes an online clustering technique, which continuosly clusters similar tweets and then classifies the clusters using Support Vector Machine algorithm.
+\textit{TEDAS} \cite{li2012tedas} detected, analysed and identified events using refined rules (e.g., keywords, hashtags) and classify messages based on content as well Twitter specific features as URLs, hashtags and mentions.
+Besides, location information is extracted using both explicit geographical coordinates and implicit geographical references in the content.
+In the same way, \citeauthor{becker2011beyond} \cite{becker2011beyond} proposed an on-line clustering technique, which continuously clusters similar tweets and then classifies the clusters using Support Vector Machine algorithm.
 Finally, events (clusters) are classified into real-world events or nonevents.
 
-Summarizing, most previous works to detect an emergency situations rely on keywords using probabilistic temporal models for specific domains (eg.
-keywords or location).
-In general terms, this approach needs knowledge about the event and a new emerging topic referring to a crisis can not identify to detect an emergency situation.
-In the case of the works about event detection based on locations, they are designed to be applying in a small area and based on historical data.
-Regardless of which approach is implemented (supervised or unsupervised), textual features are the most used attribute to characterize an event and online clustering is the most common technique to create candidate events.
+\medskip
+
+\noindent{\textbf{Summary.}} Most previous works to detect an emergency situations rely on keywords using probabilistic temporal models for specific domains (e.g., keywords or location).
+In general terms, these approaches need knowledge about the event and a new emerging topic referring to a crisis cannot be identified to detect an emergency situation.
+In the case of the works on event detection based on locations, they are designed to be applied in a small area and based on historical data.
+Regardless of which approach is implemented (supervised or unsupervised), textual features are the most used attributes to characterise an event and on-line clustering is the most common technique to create candidate events.
+
+Based on the works presented by \citeauthor{guzman2013line} \cite{guzman2013line} and \citeauthor{maldonado2017} \cite{maldonado2017}, we extend the ideas of bursty keywords and \textit{z-score} variation between fixed time-windows and we apply these proposals over locations to identify anomalies in social media activity.
+Furthermore, we do not use a set of keywords to detect specific types of events (e.g., earthquake, flood or terrorist attack).
+In contrast, our proposal is focused on detecting such events by tracking frequencies, and probability distributions of the interarrival time of the messages related to specific locations.
 
@@
 
-In order to provide a complete coverage of location-based detection of emergency situation, we propose an approach which has four stages as we can in Data Process module in Figure \ref{fig:diagram}.
-The following information provides details of each stage.
+In order to provide a complete coverage of location-based detection of emergency situations, we propose an approach which has four stages (depicted in the Data Process module in Figure \ref{fig:diagram}).
+Next, we describe each stage:
 
@@
 %To increase the number of messages related to locations, we consider tweets and retweets messages. 
-Since that we consider users as citizen sensors, we filter messages depending of native language of each country using the attribute \textit{lang} retrieved from Tweet Metadata\footnote{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object}.
-For example, if we analyze Italy, we just consider messages in italian.
-Finally, we remove user mentions, urls,  special characters and apply text tokenization .
-Also we do not remove hashtags or stopwords because some locations can be included in hashtags terms and some places can be contained stopword that can differentiate them from other similar non-related to locations terms.
+Since we consider users as citizen sensors, we filter messages depending of native language of each country using the attribute \textit{lang} retrieved from the metadata in tweets.\footnote{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object} For example, if we analyse Italy, we just consider messages in Italian.
+Finally, we remove user mentions, URLs,  special characters and apply text tokenization.
+Also we do not remove hashtags or stopwords because some locations can be included in hashtags terms and some places can contain stopwords that can differentiate them from other similar non-related to locations terms.
 
-\subsection{Signals Creation}
-The problems that we address in this stage is how to infer locations from micro-blog messages to track signal over time in the Stream and what is the lowest hierarchy level to find and assign to this signal.
+\subsection{Signal Creation}
+The problems that we address in this stage are how to infer locations from microblog messages to track signals over time in the data stream and what is the lowest hierarchy level to find and assign to these signals.
 
 \subsubsection{Geographical Hierarchy}\label{sssec:geohie}
-Using the idea of \textit{Gazetteer as a Tree} presented in \cite{yin2014pinpointing} in which each place is associated with a canonical taxonomy node, we construct our gazetteer tree based on Geonames\footnote{http://download.geonames.org/export/dump/} and Wikipedia\footnote{http://www.wikipedia.org/} for each country to analyze.
-However, in \cite{yin2014pinpointing} the gazetteer hierarchy presents four levels where the lowest level represents a specific POIs.
-In our approach, we use a subset of the gazetteer hierarchy with three levels, in which the lowest level is represented by City since a large amount of users specify their location down to the city scale \cite{hecht2011tweets}.
+Using the idea of \textit{Gazetteer as a Tree} presented in \cite{yin2014pinpointing} in which each place is associated with a canonical taxonomy node, we construct our gazetteer tree based on Geonames\footnote{http://download.geonames.org/export/dump/} and Wikipedia\footnote{http://www.wikipedia.org/} for each country to analyse.
+However, in \cite{yin2014pinpointing} the gazetteer hierarchy presents four levels where the lowest level represents a specific point of interest.
+In our approach, we use a subset of the gazetteer hierarchy with three levels, in which the lowest level is represented by a city since a large amount of users specify their location down to this level \cite{hecht2011tweets}.
 For example, if we have the \textit{city:Manchester}, we associate this location with \textit{region-state:North West} and also with \textit{country:England}.
@@
 
-\subsubsection{Finding Locations in Micro-Blog Metadata}
+\subsubsection{Finding Locations in Microblog Metadata}
 
-The structure of the Tweet Metadata allows to obtain information about the message and the user who shares a message.
+The structure of the Tweet metadata obtains information about the message and the user who shares a message.
 Considering the geographical hierarchy explained in the previous section, we search these locations on different levels of metadata and create one signal for each as following:
@@
 	\item Tweet Text - User Location: location is mentioned in the attribute \texttt{text} of tweet object and also location is mentioned in the attribute \texttt{location} inside the \texttt{user object}.
-This means that the location is mentioned on the body of the message and the user who share message has the same location in his profile.
-In this case, tweet text and user location can be different in the smallest hierarchy but in the highest level can be equal.
+This means that the location is mentioned in the body of the message and the user who shares message has the same location in his profile.
+In this case, tweet text and user location can be different in the smallest hierarchy, but in the highest level can be equal.
 \end{itemize}
 
-In this way, mixing geographical hierarchy and locations in micro-blog metadata, we create $NxM$ signals where $N$ is the number of locations obtained by gazetteer tree and $M$ is the number of metadata-levels extracted from Tweet Object.
-In instance, we create a signal for \textit{city:Manchester} and we find this hierarchy in \textit{metadata:Tweet Text} and also in \textit{metadata:User Location}.
+In this way, mixing geographical hierarchy and locations in microblog metadata, we create $NxM$ signals where $N$ is the number of locations obtained by gazetteer tree and $M$ is the number of metadata-levels extracted from Tweet Object.
+For instance, we create a signal for \textit{city:Manchester} and we find this hierarchy in \textit{metadata:Tweet Text} and also in \textit{metadata:User Location}.
 That means that we track the mention of \textit{city:Manchester} at the level of the body of message and at the level of the location of the user profile individually.
@@
 \subsection{Time-Windows}
-In this stage we address the problems of how to divide and determine the time-windows size to detect a new emergency situation and what features by time-window allow it.
+In this stage we address the problems of how to divide and determine the time-window size to detect a new emergency situation and what features by the time-window allow it.
 
@@
 
-According to \cite{guzman2013line}: ``If the window size is to small, the ocurrence of empty windows for a term increases, making the noise rate increate and frequency rate tend to zero.
+According to \citeauthor{guzman2013line} \cite{guzman2013line}: ``If the window size is too small, the ocurrence of empty windows for a term increases, making the noise rate increase and frequency rate tend towards zero.
 On the other hand, if the window size is too large, the stability of the signal becomes constant and bursty keyword detection is delayed''.
@@
 %\end{itemize}
-\subsubsection{Inter-arrival Time}
+\subsubsection{Interarrival Time}
 
-To characterize the urgency of the messages during a time-window, we use our previous results \cite{DBLP:conf/fdia/Sarmiento17} to compute the \textit{Inter-arrival Time} which is defined as $d_{i} = t_{i+1} - t_{i}$, where $d_{i}$ denotes the difference between two consecutive social media messages $i$ and $i+1$ that arrived in moments $t_{i}$ and $t_{i+1}$ respectively.
-Using this definition and according to \cite{kalyanam2016prediction}, high-activity events have a high-frequency in the first bins represented by values $d_{i} \approx 0$.
+To characterize the urgency of the messages during a time-window, we use our previous results \cite{DBLP:conf/fdia/Sarmiento17} to compute the \textit{interarrival time} which is defined as $d_{i} = t_{i+1} - t_{i}$, where $d_{i}$ denotes the difference between two consecutive social media messages $i$ and $i+1$ that arrived in moments $t_{i}$ and $t_{i+1}$ respectively.
+Using this definition and according the work of \citeauthor{kalyanam2016prediction} \cite{kalyanam2016prediction}, high-activity events have a high-frequency in the first bins represented by values $d_{i} \approx 0$.
 
-In order to quantify a high-frequency in very small values of $d_{i}$, we compute the measures \textit{Skewness} and \textit{Kurtosis}, which represent the asymmetry and the tailedness of the shape of probability distribution respectively \cite{mardia1970measures}.
-Finally, we apply the equation \ref{eq:1} over \textit{Skewness} and \textit{Kurtosis} to calculate variation based on previous values.
+To quantify a high-frequency in very small values of $d_{i}$, we compute the measures \textit{skewness} and \textit{kurtosis}, which represent the asymmetry and the tailedness of the shape of probability distribution respectively \cite{mardia1970measures}.
+Finally, we apply the equation \ref{eq:1} over \textit{skewness} and \textit{kurtosis} to calculate variation based on previous values.
 
@@
 
-An emergency situation that affects and mobilizes response in a small area is defined as \textit{Focalized}, while a disaster with a large geographic impact is defined as \textit{Diffused} \cite{olteanu2015expect}.
+An emergency situation that affects and mobilizes response in a small area is defined as \textit{focalized}, while a disaster with a large geographic impact is defined as \textit{diffused} \cite{olteanu2015expect}.
 Using this definition, we extend this concept to represent neighborhoods between locations obtained from section \ref{sssec:geohie}.
-For that purpose, we create an \textit{Adjancency Matrix} $M$, where $M_{i,j} = 1$ represents if two locations are geographically connected and $M_{i,j} = 0$ if they are not connected.
-For instance, if an event is diffused (eg.
-earthquake), the detection should be in adjacent-locations independentely of metadata-level.
-In the other hand, if an event is focalized (eg.
-terrorist attack), just one location should be detected but in different metadata-levels simultaneously.
+For that purpose, we create an \textit{adjancency matrix} $M$, where $M_{i,j} = 1$ represents if two locations are geographically connected and $M_{i,j} = 0$ if they are not connected.
+For instance, if an event is diffused (e.g., earthquake), the detection should be in adjacent-locations independentely of metadata-level.
+On the other hand, if an event is focalized (e.g., terrorist attack), just one location should be detected but in different metadata-levels simultaneously.
 
@@
 
-We analyze five earthquakes with magnitudes between $5.5Mw$ and $7.6Mw$\footnote{Mw: the moment magnitude scale}, ocurring in Italian-speaking and Spanish-speaking countries between October 2016 and April 2017 (Table \ref{tab:eqs}).
-For that purpose, we collect $20$ millions of messages 12-hour before and after the emergency situation events.
+We analyse five earthquakes with magnitudes between $5.5Mw$ and $7.6Mw$\footnote{Mw: the moment magnitude scale}, ocurring in Italian-speaking and Spanish-speaking countries between October 2016 and April 2017 (Table \ref{tab:eqs}).
+For that purpose, we collect $20$ million of messages 12-hours before and after the emergency situation events.
 
-According our proposal, we create both the gazetteer hierarchies\footnote{http://users.dcc.uchile.cl/$\sim$hsarmien/gazetteer.html} for each country and construct the signals based on each hierarchy and metadata-level.
+According our proposal we create both the gazetteer hierarchies\footnote{http://users.dcc.uchile.cl/$\sim$hsarmien/gazetteer.html} for each country and construct the signals based on each hierarchy and metadata-level.
 As a result of the number of messages of each signal (Table \ref{tab:message_signal}), we discard all signals related to city hierarchy since a great amount of small cities have zero frequency in a normal situation unlike to capital or metropolitan cities.
@@
 National Institute of Geophysics and Volcanology in Italy\footnote{http://www.ingv.it/it/}.
-With the purpose of to labeling a time-window as positive class (detection), we set as detection those time-windows with positive variation in frequency, skewness and kurtosis respect to the normalization of the previous values.
-Moreover and according to (Figure \ref{fig:labeled}), we include the three next time-windows after the event to compensate the imbalance between classes given that later to these number of time-windows, the variation in the features decrease.
+With the purpose of labeling a time-window as positive class (detection), we set as detection those time-windows with positive variation in frequency, skewness and kurtosis with respect to the normalization of the previous values.
+Moreover and according to (Figure \ref{fig:labeled}), we include the three next time-windows after the event to compensate the imbalance between classes given that after these number of time-windows, the variation in the features decrease.
 
@@
 
-In order to classifier messages, we employ traditional binary classifier Support Vector Machine \textit{(SVM)}.
-As a result of the analyzed data scattering (Figure \ref{fig:scatter}), we separate country and state in different datasets and set both kernels and classification parameters independentely.
-On the one hand, \textit{country} classifier uses a polynomial kernel and strict-parameters for gamma, cost and weights since that a great amount of messages are included in country hierarchy as an effect of the minor hierarchies.
+To classify messages we employed traditional binary classifier Support Vector Machine \textit{(SVM)}.
+As a result of the analysed data scattering (Figure \ref{fig:scatter}), we separated country and state in different datasets and set both kernels and classification parameters independentely.
+On the one hand, \textit{country} classifier uses a polynomial kernel and strict-parameters for gamma, cost and weights since a great amount of messages are included in country hierarchy as an effect of the minor hierarchies.
 On the other hand, \textit{state/region} classifier uses a linear kernel with soft-weights and cost.
@@
 Given that an emergency situation is not an usual event, we have an highly unbalanced data respect to the classes after labeled ($1\approx2\%$ of positive class corresponding to \textit{detection}).
-Therefore, we use \textit{under-sampling} \cite{lunardon2014rose} over country and state datasets increasing our possitive class to $15\approx18\%$.
-Additionally to validate our model, we use \textit{5-fold cross-validation} where one earthquake dataset is used as testing and the reamining earthquakes dataset as training.
+Therefore, we used \textit{under-sampling} \cite{lunardon2014rose} over country and state datasets increasing our possitive class to $15\approx18\%$.
+Additionally, to validate our model, we used \textit{5-fold cross-validation} where one earthquake dataset is used as testing and the remaining earthquakes dataset as training.
 
@@
 Our first analysis is just considering the hierarchies as isolated detections.
-On the top of the Table \ref{tab:perfomance}, we show the results considering only the prediction over each label in our datasets.
-As noted above, the assignation from the lowest level (\textit{city}) to the highest (\textit{country}) in the gazetteer hierarchy generates high frequency of messages which cause multiples \textit{bursts} in our country signal for non  emergency situations.
+The top of the Table \ref{tab:perfomance} shows the results considers only the prediction over each label in our datasets.
+As noted above, the assignation from the lowest level (\textit{city}) to the highest (\textit{country}) in the gazetteer hierarchy generated high frequency of messages which cause multiple \textit{bursts} in our country signal for non  emergency situations.
 This concept can explain the values of Precision (\textit{P}) and \textit{FPR}.
@@
 For this analysis, we search the time-windows for each hierarchy where the all metadata-levels are well classified with correct class.
-According to the results shown on the middle of the Table \ref{tab:perfomance}, when we analyze country and state independentely the values of Precision, F1 and FPR have worst values than the analysis by label .
+According to the results shown on the middle of the Table \ref{tab:perfomance}, when we analyse country and state independentely the values of Precision, F1 and FPR have worst values than the analysis by label .
 \begin{table}
@@
 \subsubsection{Dependent Analysis of Hierarchies}
-Our second analysis is considering the hierarchies as non-isolated detections.
-In the results explained above, we consider country and state hierarchy independently, which is not a correct analysis because an emergency situation affect to states and country at the same time.
-For this reason, we inspect the time-windows where all metadata-level for country and state hierarchy have a correct detection simultaneously.
+Our second analysis considered the hierarchies as non-isolated detections.
+In the results explained above, we considered country and state hierarchy independently, which is not a correct analysis because an emergency situation affects states and country at the same time.
+For this reason, we inspected the time-windows where all metadata-level for country and state hierarchy have a correct detection simultaneously.
 The results are shown in the row \textit{Country-State} in the Table \ref{tab:perfomance}.
-In contrast to the independent analysis of country and state, we improve the Precision, F1 and FPR values as a consequence of a smaller amount of the time-windows related to non emergency situations are assigned as detection.
+In contrast to the independent analysis of country and state, we improved the Precision, F1 and FPR values as a consequence of a smaller amount of the time-windows related to non emergency situations are assigned as detection.
 However, when we see the value obtained for FPR ($FPR = 0.03$), this rate represents an incorrect number of time-windows assigned as detection equal to $23$.
@@
 \subsubsection{Geopraphic Spread Analysis}\label{sssec:geospreadanalysis}
-
-Our third analysis is considering the hierarchies as non-isolated detections and applying the Geographic Spread (G.S.).
-Using the \textit{Adjacency Matrix} to represent neighborhoods between regions/states, we consider as a correct detection those time-windows where the state/s classified as detection are defined as \textit{Focalized} or \textit{Diffused} and exist dependency between hierarchies.
+Our third analysis considered the hierarchies as non-isolated detections and applies the Geographic Spread (G.S.).
+Using the \textit{Adjacency Matrix} to represent neighborhoods between regions/states, we considered as a correct detection those time-windows where the state/s classified as detection are defined as \textit{Focalized} or \textit{Diffused} and exist dependency between hierarchies.
 
@@
 		U.K. Elections& 5 & - & - & - & \small{(theresa, may), (vote, labour), (van, dijk)}\\
-		Adele Live at Wembley& 9& 7&2&-& \small{(elland, road), (new, times), (phil, jackson)}\\
-		England vs Eslovenia Soccer Match & 4& 4 & 0 & - & \small{(simon, brodkin), (join, us), (theresa, may)}\\
-		Metallica Live at London& 4 & 4 & 0 & - & \small{(always, said), (chance, win), (carabao, cup)}\\
+		Adele Live in Wembley& 9& 7&2&-& \small{(elland, road), (new, times), (phil, jackson)}\\
+		England vs Slovenia Soccer Match & 4& 4 & 0 & - & \small{(simon, brodkin), (join, us), (theresa, may)}\\
+		Metallica Live in London& 4 & 4 & 0 & - & \small{(always, said), (chance, win), (carabao, cup)}\\
 		\bottomrule
@@
 		U.K. Elections& 0 & - & - & - & \hfill \break\\
-		Adele Live at Wembley& 0& 0&0&-& \hfill \break\\
-		England vs Eslovenia Soccer Match & 1& 1 & 0 & - & \small{(per, day), (menswear, sample), (closed, roads)}\\
-		Metallica Live at London& 2 & 2 & 0 & - & \small{(happy, birthday), (chance, win), (always, said)}\\
+		Adele Live in Wembley& 0& 0&0&-& \hfill \break\\
+		England vs Slovenia Soccer Match & 1& 1 & 0 & - & \small{(per, day), (menswear, sample), (closed, roads)}\\
+		Metallica Live in London& 2 & 2 & 0 & - & \small{(happy, birthday), (chance, win), (always, said)}\\
 		\bottomrule
@@
 \end{table*}
+
 In addition to the results of the dependency analysis explained above, we see that a large amount of time-windows for country hierarchy ($\approx 82\%$) have more than one metadata-level when exist a correct detection.
@@
 
-Considering the geographic spread by states and the number of metadata-levels by country hierarchy, we analyze the results shown on the bottom of the Table \ref{tab:perfomance}.
-On the one hand, the row with value equal to \textit{Country(2)-State with G.S.} represent the detection when we consider at least two metadata-levels for the country hierarchy and the geographic spread for states.
-In contrast to the the previous analyzes, we improve the values of the Precision, F1 and FPR.
-The last metrics is very important because there are no time-windows incorrectly assigned as emergency situations.
+Considering the geographic spread by states and the number of metadata-levels by country hierarchy, we analysed the results shown on the bottom of the Table \ref{tab:perfomance}.
+On the one hand, the row with value equal to \textit{Country(2)-State with G.S.} represents the detection when we considered at least two metadata-levels for the country hierarchy and the geographic spread for states.
+In contrast to the the previous analyses, we improved the values of the Precision, F1 and FPR.
+The last metrics are very important because there are no time-windows incorrectly assigned as emergency situations.
 Consequently, the Recall values decrease which means that our method remove some time-windows classified as detection.
@@
 
-On the other hand, \textit{Country(3)-State with G.S.} represent the detection when we consider three metadata-levels for country hierarchy and the geographic spread for states.
-Similar to  \textit{Country(2)-State with G.S.}, we improve the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with a average delay equal to $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the first detection.
+On the other hand, \textit{Country(3)-State with G.S.} represent the detection when we considered three metadata-levels for country hierarchy and the geographic spread for states.
+Similar to  \textit{Country(2)-State with G.S.}, we improved the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with a average delay equal to $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the first detection.
 
 \subsection{Online Evaluation}
-For our evaluation in the Twitter Public Stream, we training a classifier with five earthquakes identified in our ground truth.
-Futhermore, our evaluation dataset is formed by eight different events that ocurred in England between December 2016 and October 2017.
-For each event we consider the full-day in which they ocurred.
-The main goals of this evaluation is to know the capacity of our method to detect emergency situations and discard non-related to emergency situations events that involve references of locations.
+For our evaluation in the Twitter Public Stream, we trained classifier with five earthquakes identified in our ground truth.
+Furthermore, our evaluation dataset is formed by eight different events that occurred in England between December 2016 and October 2017.
+For each event we considered the full-day in which they ocurred.
+The main goals of this evaluation is to know the capacity of our method to detect emergency situations and discard those non-related to emergency events that involve location references.
 Geographic spread analysis is used to evaluate our method because decrease the number of false positives detection.
@@
 
-As can be appreciated on Table \ref{tab:online1} and Table \ref{tab:online2}, we study three terrorist attacks and five high-impact real-world events related to soccer matches, music concerts and political elections.
-In the case of Premier League Soccer Matches and U.K Elections, we can not identify the beginning of the event, since that in the first one there are many soccer matches during the analyzed day and in the second one there is no a specific start time.
-In order to know the topics when our method detect an event, we compute the Top 3 Bigrams in the detected time-windows.
-Also, we calculate the delay time just for emergency situations events.
+As can be noted on Table \ref{tab:online1} and Table \ref{tab:online2}, we study three terrorist attacks and five high-impact real-world events related to soccer matches, music concerts and political elections.
+In the case of Premier League Soccer Matches and U.K Elections, we can not identify the beginning of the event, since in the first one there are many soccer matches during the analysed day and in the second one there is no a specific start time.
+In order to know the topics when our method detects an event, we computed the Top 3 Bigrams in the detected time-windows.
+Also, we calculated the delay time just for emergency events.
 
@@
 However, the London Terrorist Attack has $50\%$ of the detected time-windows after the event, which means that there are seven time-windows non-related to emergency situations.
-Besides to the crisis situations analysis, we also study the number of detected time-windows in non-related to emergency situation events.
+Besides the crisis situations analysis, we also study the number of detected time-windows in non-related to emergency situation events.
 In the same way, we have a large amount of misclassified time-windows that do not represent crisis situations as we can see in the Top 3 Bigrams for each non-related to event.
@@
 On the other hand, the second evaluation \textit{Country(3)-State with G.S.} decreases the number non-related to emergency situations events detected as crisis situations.
-We can see three time-windows in two events detected as emergency situations (England vs Eslovenia, and Metallica Live at London).
+We can see three time-windows in two events detected as emergency situations (England vs Eslovenia, and Metallica Live in London).
 In these cases, the time-windows are detected before the event and corresponding a non emergency situations according to the bigrams.
-Furthermore, when we analyzed the number of the detected emergency situations, two-third ($66\%$) of the events are detected correctly with average delay time equal to $30.3$ minutes .
-In the case of London Terrorist Attack, our method detects one time-window before the event but the bigrams describe that the detection do not correspond to crisis situation.
+Furthermore, when we analysed the number of the detected emergency situations, two-thirds ($66\%$) of the events are detected correctly with average delay time equal to $30.3$ minutes.
+In the case of London Terrorist Attack, our method detects one time-window before the event but the bigrams describe that the detections do not correspond to crisis situations.
 
@@
 	\includegraphics[width=\columnwidth]{img/delay.png}
-	\caption{Delay time and number of locations in the first detection for diffused and focalized emergency situacions.}
+	\caption{Delay time and number of locations in the first detection for diffused and focalized emergency situations.}
 	\label{fig:delay}
@@
 Our findings suggest that there is evidence to detect an emergency situation based on anomaly frequency of messages that contain locations for a specific country.
-Indeed, our method based on the number of metadata-levels by country hierarchy and geographic spread by state, detect $80\%$ of the events related to emergency situations as we could demostrate in our ground truth.
-Also, our method is independent of the textual features because we apply the model over different languages as spanish, italian and english.
-Furthermore, we test our model in different types of events such as earthquakes and terrorist attacks, and also on different magnitudes (in the case of earthquakes) and number of affected people (e.g: Manchester Terrorist Attack vs Westminster Terrorist Attack).
+Indeed, our method based on the number of metadata-levels by country hierarchy and geographic spread by state, detects a $80\%$ of the events related to emergency situations as we could demonstrate in our ground truth.
+Also, our method is independent of the textual features because we apply the model over different languages as Spanish, Italian and English.
+Furthermore, we test our model in different types of events such as earthquakes and terrorist attacks, and also on different magnitudes (in the case of earthquakes) and number of affected people (e.g., Manchester Terrorist Attack vs Westminster Terrorist Attack).
 
-However, when we apply our method in the online evaluation , we detect $66\%$ of the emergency situations that affected to England.
-This can explain because the signal are noise but many reason: the number of active users in United Kingdom\footnote{http://www.statista.com/statistics/242606/number-of-active-twitter-users-in-selected-countries/ visited on January 2018} which can affect the anomaly frequency of message since that exist a high daily average activity of the messages; similar locations in other countries (York $\approx$ New York); and soccer teams with names of cities (Manchester United, Liverpool).
-These issues also can affect the number of false positive detection in which in the case of England was $30\%$ of the non-related to emergency situations events.
+However, when we apply our method in the on-line evaluation, we detect $66\%$ of the emergency situations that affected England.
+This explains that the signals, and for various reasons: the number of active users in United Kingdom\footnote{http://www.statista.com/statistics/242606/number-of-active-twitter-users-in-selected-countries/ visited on January 2018} which can affect the anomaly frequency of the messages since there exists a high daily average activity of the messages; similar locations in other countries (York $\approx$ New York); and soccer teams with names of cities (Manchester United, Liverpool).
+These issues also can affect the number of false positive detection in which in the case of England was $30\%$ of the non-related to emergency events.
 
-Reggarding the geographic spread where we define an emergency situation as diffused or focalized, we find some evidences that differentiate them.
+Regarding the geographic spread where we define an emergency situation as diffused or focalized, we find some evidence that differentiates them.
 In the case of diffused events, the delay time of the our first detection was less than $12$ minutes and in focalized events was greater than $30$ minutes (Figure \ref{fig:delay}).
-This can be explain given that, in diffused events such as earthquakes, a large amount of people are affected (thousand or millions) at the same time by an event which generate a collective reaction in social media in the locations where the event impacted.
+This explains that, in diffused events such as earthquakes, a high number of people are affected (thousands or millions) at the same time by an event which generates a collective reaction in social media in the locations where the event impacted.
 In Figure \ref{fig:delay}, we can see that earthquakes have at least two detected locations in the first detection (except Italy EQ2).
-In contrast, focalized events have less amount of eyewitness (hundred or thousands) then when the users share messages in social media, the frequency not affect the average daily message of the country in the first minutes.
-This can be explain en Figure \ref{fig:delay} where the terrorist attacks have just 1 detected locations in the first detection.
+In contrast, focalized events have less amount of eyewitness (hundreds or thousands) then when the users share messages in social media, the frequency does not affect the average daily message of the country in the first minutes.
+This can be explained in Figure \ref{fig:delay} where the terrorist attacks have just 1 detected location in the first detection.
 
-Additionally, the delay time can be different according to many reason: datetime of the event (for example, during the early hours), little differences with the end of the current time-window, type of the affected locations (rural, urban cities) and the number of active users by locations.
+Additionally, the delay time can be different for many reasons: datetime of the event (for example, during the early hours), few differences with the end of the current time-window, type of the affected locations (rural, urban cities) and the number of active users by locations.
 
@@
 This approach is independent of the textual features and can be used in different types of events and languages.
-We show that the users act self-organized in the affected locations like citizen sensors when an emergency situation occur.
-We furthermore have presented an analysis of geographic spread for different types of events, allowing characterize them.
-However, our experiment consider just a small portion of emergency situations which is not representive for all type of crisis situations according either the hazard type (natural or human-induced), temporal development (instantaneous or progressive) or geographic spread (diffused or focalized).
+We show that the users act as self-organized in the affected locations like citizen sensors when an emergency situation occurs.
+We furthermore have presented an analysis of geographic spread for different types of events that can be categorized.
+However, our experiment considers just a small portion of emergency situations, which is not representive for all types of crisis situations according to either the hazard type (natural or human-induced), temporal development (instantaneous or progressive) or geographic spread (diffused or focalized).
 
-There are many things that can be improved our results.
+There are many things that can improve our results.
 We will add Point of Interest to our gazetteer tree to increase the frequency by time-windows in each hierarchy.
 Furthermore, we will add more non-textual features as number of retweets and tweets, unique locations detected and special locations.
-We also plan study the relevance of the different metadata-levels and assign weights for each.
-Finally, we will create a web application to visualize in real-time events.
+We also plan to study the relevance of the different metadata-levels and assign weights for each.
+Finally, we will create a web application to visualize events in real-time.
 
-\begin{acks}
+%\begin{acks}
 
-\end{acks}
+%\end{acks}
 

### 2018-02-22 paper_OLD.tex
add barbara's review as introduction to some sections and sort experiments and results

--- 2018-02-05 paper.tex
@@
 We furthermore have presented an analysis of geographic spread for different types of events that can be categorized.
-However, our experiment considers just a small portion of emergency situations, which is not representive for all types of crisis situations according to either the hazard type (natural or human-induced), temporal development (instantaneous or progressive) or geographic spread (diffused or focalized).
+However, our experiment considers just a small portion of emergency situations, which is not representative for all types of crisis situations according to either the hazard type (natural or human-induced), temporal development (instantaneous or progressive) or geographic spread (diffused or focalized).
 

### 2018-02-23 paper.tex
add introduction to several sections

@@
 
-\author{Sergio Campos}
+\author{Jaime Campos}
 %\authornote{This author is the
@@
 %
-Most methods for detecting emergency situations using Twitter rely on identifying features within messages that contain domain-specific keywords.
-However, keyword-based methods require models to be trained on historical data of specific domains, in multiple languages, and for different types of events (e.g., earthquakes, floods, wildfires, etc.).
+In general, existing methods for automatically detecting emergency situations using Twitter rely on features based on domain-specific keywords found in messages.
+This type of keyword-based methods usually require training on domain-specific labeled data, using multiple languages, and for different types of events (e.g., earthquakes, floods, wildfires, etc.).
 %
@@
 %
-Nevertheless, collective mentions of keywords are not the only type of self-organizing phenomena that may arise when a real-world extreme situation occurs.
+However, collective mentions of certain keywords are not the only type of self-organizing phenomena that may arise in social media when a real-world extreme situation occurs.
 %
-Just as physical sensors become activated when stimulated, localized citizen sensors will also react in a similar manner.  
+Just as nearby physical sensors become activated when stimulated, localized citizen sensors (i.e., users) will also react in a similar manner.  
 %
-To leverage this information, we propose to use self-organized geolocation related activity to identify emergency situations.
+To leverage this information, we propose to use self-organized activity related to geolocations to identify emergency situations.
 %
@@
 Instead, microblogging has played a critical role over the last fifteen years allowing users to share real-time information from people local to the incident, such as status updates, casualties, damages and alerts \cite{kumar2011tweettracker,imran2013extracting,stowe2016identifying,reuterfifteen}.
-For this reason, researchers have studied user behavior during these events to detect, summarise and classify messages with the goal of helping authorities, and the general public, with situational awareness to provide fast and conscientious responses during crisis situations.
+For this reason, researchers have studied user behavior during these events to detect, summarize and classify messages with the goal of helping authorities, and the general public, with situational awareness to provide fast and conscientious responses during crisis situations.
 
@@
 This is because, as with most social media conversations, messages are often overridden with irrelevant and redundant noise.
-In current works \cite{kumar2011tweettracker,caragea2011classifying,imranaidr2014,maldonado2017}, these tasks are solved with methods that rely on keyword based filters on the \textit{Twitter Public Streaming API}.\footnote{https://developer.twitter.com/en/docs/tweets/filter-realtime/overview} The problem with keyword-based methods is the need to train in specific domains for different type of events.
-For instance, \citeauthor{olteanu2014} \cite{olteanu2014} generate a set of keywords based on different datasets.
-However, this is not sufficient for cases in which specific and previously unseen terms arise for a particular event (e.g., \textit{\#eqnz} for Earthquakes in New Zealand, or \textit{\#pabloph} for Typhoon Pablo in Philippines) \cite{potts2011tweeting,bruns2012local,karimi2013classifying}.
-Furthermore, these sets of keywords are commonly obtained a particular language and will not work in others.
+In current works \cite{kumar2011tweettracker,caragea2011classifying,imranaidr2014,maldonado2017}, these tasks are solved with methods that rely on keyword based filters on the \textit{Twitter Public Streaming API}.\footnote{https://developer.twitter.com/en/docs/tweets/filter-realtime/overview} However, a drawback of keyword-based methods is that they require to train for each specific domain, and for different type of events.
+For instance, \citeauthor{olteanu2014} \cite{olteanu2014} generated a set of keywords for event detection based on different datasets.
+However, these keywords are not sufficient for cases in which specific, and previously unseen, terms arise for particular events (e.g., the hashtags: \textit{\#eqnz}, used for Earthquakes in New Zealand, and \textit{\#pabloph}, used for Typhoon Pablo in Philippines) \cite{potts2011tweeting,bruns2012local,karimi2013classifying}.
+Furthermore, domain-specific keywords are commonly extracted using a particular language and do not apply to others.
 
-Additionally, several works related to crisis situations show a strong relationship between a type of event (what) and the spatiotemporal dimensions (when and where).
-For example, it has been observed that top trends during earthquakes are related to location mentions \cite{mendoza2010twitter}, also, that there is a strong relationship between the proximity to a hurricane path and hurricane-related social media activity \cite{kryvasheyeu2016rapid}.
+In addition, prior work related to crisis situations in social media, show a strong relationship between the type of event (the {\em what}) and spatio-temporal dimensions (the {\em when} and {\em where}).
+For example, the top-trends during earthquakes are usually related to location mentions \cite{mendoza2010twitter}, and that there is a strong relationship between the proximity to a hurricane path and hurricane-related social media activity \cite{kryvasheyeu2016rapid}.
 Other works have addressed the extraction of locations and points of interest during floods \cite{lingad2013location}, as well as mixing geographic information system (GIS) information with geo-tagged messages to improve disaster mapping and real-time event tracking \cite{huang2015disastermapper}.
 
-Based on previous works, the question we address here is: \textit{Is there evidence to detect an emergency situation based on anomalies in messages related to contain locations?}.
+In order to complement prior findings, we propose to address the following research question: \textit{Can we detect automatically emergency situations, based solely on location-related information found in user messages?}.
 
-Here, we propose a method based on recurring references to country level locations in message metadata.
+We propose a method based on recurring references to country-level locations in message metadata.
 For this task, we create a gazetteer tree based on the hierarchy of the specific country, divide the messages into fixed time-windows and compute the frequency and the probability distributions of the interarrival time of the messages for each geographical hierarchy.
-To detect an emergency situation, we train a SVM classifier by each hierarchy and apply a geographic spread to filter false positives detection considering that an emergency situation can be \textit{Focalized} or \textit{Diffused}.
+To detect an emergency situation, we train a SVM classifier in each hierarchy and apply a geographic spread to filter false positives detection considering that an emergency situation can be either \textit{focalized} (i.e., that affects a reduced area) or \textit{diffused} (i.e., that affects a large area).
 
 Our main contribution is to create a methodology that detects instantaneous emergency situations just by using the location related information for a specific country.
-Furthermore, our classifiers do not depend on language to detect a new event since they do not use textual features as input.
-Moreover, we characterize crisis situations that affect small or large geographical areas.
+Furthermore, our approach does not depend on language to detect a new event since it does not use textual features as input.
+In addition, we characterize focalized and diffused crisis situations.
 
 The paper is organized as follows: we first introduce an overview of relevant literature related to emergency situation and event detection in social media.
-Next we present a complete description of our proposal divide in four parts.
-Thereafter, we provide a full description of our dataset and later we summarise our experimental validation over the ground truth and on-line evaluation.
+Next we present a complete description of our proposal.
+Then, we provide a full description of our dataset and later we summarize our experimental validation over the ground truth and on-line evaluation.
 Finally, we deliver our discussions, conclusions and future work.
@@
 To detect such events, they identify a group of tweets that describe a particular theme, which are generated within a short time frame and a same geographic area.
-Similarly, \citeauthor*{unankard2015emerging} \cite{unankard2015emerging} proposed an approach for early detection of emerging hotspot events in social networks with location sensivity.
+Similarly, \citeauthor*{unankard2015emerging} \cite{unankard2015emerging} proposed an approach for early detection of emerging hotspot events in social networks with location sensitivity.
 In this work, the authors identified strong correlations between user locations and event locations when detecting emerging events using content similarity between clusters.
@@
 Supervised approaches have also been proposed.
-\textit{TEDAS} \cite{li2012tedas} detected, analysed and identified events using refined rules (e.g., keywords, hashtags) and classify messages based on content as well Twitter specific features as URLs, hashtags and mentions.
+\textit{TEDAS} \cite{li2012tedas} detected, analyzed and identified events using refined rules (e.g., keywords, hashtags) and classify messages based on content as well Twitter specific features as URLs, hashtags and mentions.
 Besides, location information is extracted using both explicit geographical coordinates and implicit geographical references in the content.
@@
 
-\noindent{\textbf{Summary.}} Most previous works to detect an emergency situations rely on keywords using probabilistic temporal models for specific domains (e.g., keywords or location).
-In general terms, these approaches need knowledge about the event and a new emerging topic referring to a crisis cannot be identified to detect an emergency situation.
-In the case of the works on event detection based on locations, they are designed to be applied in a small area and based on historical data.
-Regardless of which approach is implemented (supervised or unsupervised), textual features are the most used attributes to characterise an event and on-line clustering is the most common technique to create candidate events.
+\noindent{\textbf{Summary.}} Most of the existing work for detecting emergency situations, rely on keywords using probabilistic temporal models for specific domains (e.g., using keywords or locations).
+In general, these approaches require background knowledge about the event, and therefore, they will not identify new types of emerging crisis situations.
+In the case of existing approaches for event detection based on locations, these have been designed for small areas and are based on historical data.
+Regardless of the type of approach used (supervised or unsupervised), events are always characterized based on textual features, and on-line clustering is the most common technique used to create candidate events.
 
 Based on the works presented by \citeauthor{guzman2013line} \cite{guzman2013line} and \citeauthor{maldonado2017} \cite{maldonado2017}, we extend the ideas of bursty keywords and \textit{z-score} variation between fixed time-windows and we apply these proposals over locations to identify anomalies in social media activity.
-Furthermore, we do not use a set of keywords to detect specific types of events (e.g., earthquake, flood or terrorist attack).
-In contrast, our proposal is focused on detecting such events by tracking frequencies, and probability distributions of the interarrival time of the messages related to specific locations.
+Furthermore, we do not use a set of keywords to detect specific types of events.
+In contrast, our proposal is focused on detecting bursty activity related to such events by tracking frequencies, and probability distributions of the interarrival time of the messages in specific locations.
 
@@
 
-In order to provide a complete coverage of location-based detection of emergency situations, we propose an approach which has four stages (depicted in the Data Process module in Figure \ref{fig:diagram}).
+Our focus in this proposal is to detect an emergency situation based on identifying anomalies in social media activity related to locations.
+In this way, the main task is to extract locations from messages by reducing the noise and irrelevant information.
+For this reason, data is pre-processed to allow a better analysis.
+
+For location extraction task, we used a geographical dictionary (also known as gazetteer) to create a geographic hierarchy for a specific country.
+To understand the effect of an emergency situation inside to the country, we created signals in different levels of the geographic hierarchy and the tweet metadata.
+Afterwards, signals are divided into fixed time-windows and non textual features are computed for each of these.
+Finally and with the goals of filtering false positives detection, we created a geographic spread based on the proximity between locations by using an adjacency matrix.
+
+In order to provide a complete coverage of location-based detection of emergency situations, we divided our approach into four stages (depicted in the ``data process'' module in Figure \ref{fig:diagram}).
 Next, we describe each stage:
@@
 	\centering
-	\includegraphics[width=\columnwidth]{img/diagram2.png}
+	\includegraphics[width=\columnwidth]{img/diagram4.png}
 	\caption{Key components of the proposed approach.}
@@
 
-%To increase the number of messages related to locations, we consider tweets and retweets messages. 
-Since we consider users as citizen sensors, we filter messages depending of native language of each country using the attribute \textit{lang} retrieved from the metadata in tweets.\footnote{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object} For example, if we analyse Italy, we just consider messages in Italian.
-Finally, we remove user mentions, URLs,  special characters and apply text tokenization.
-Also we do not remove hashtags or stopwords because some locations can be included in hashtags terms and some places can contain stopwords that can differentiate them from other similar non-related to locations terms.
+Our focus is on localized bursty user activity.
+Therefore, initially, we filter messages according to the most common language used in each country (using the attribute \textit{lang} in tweet metadata\footnote{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object}).
+For example, when analyzing Italy, we only take into account messages written in Italian.
+This helps to filter noise of unrelated messages.
+We also remove user mentions, URLs,  special characters, and apply text tokenization.
+We do not remove hashtags nor stopwords, because some locations can be included as hashtags, and some location names contain stopwords, which differentiate them from other locations or other terms.
 
 \subsection{Signal Creation}
-The problems that we address in this stage are how to infer locations from microblog messages to track signals over time in the data stream and what is the lowest hierarchy level to find and assign to these signals.
+We create a set of discrete-time signals for each location, which indicates each time that a message related to a specific location was posted.
+In order to explain the effect of an emergency situation in a local and national scope, we use the lowest possible geographical hierarchy level available with the aim of comparing the impact in the highest level.
+Furthermore, we study the anomalies in different metadata levels to understand how locations are shared in Twitter, for instance, either based in the locations setted by users in their profile or sharing location in their messages.
+
+%We use the lowest possible geographical hierarchy level available. 
+% described next.
+%The problems that we address in this stage are how to infer locations from microblog messages to track signals over time in the data stream and what is the lowest hierarchy level to find and assign to these signals.
 
 \subsubsection{Geographical Hierarchy}\label{sssec:geohie}
-Using the idea of \textit{Gazetteer as a Tree} presented in \cite{yin2014pinpointing} in which each place is associated with a canonical taxonomy node, we construct our gazetteer tree based on Geonames\footnote{http://download.geonames.org/export/dump/} and Wikipedia\footnote{http://www.wikipedia.org/} for each country to analyse.
+We use the idea of \textit{gazetteer as a tree} presented in \cite{yin2014pinpointing} in which each place is associated with a canonical taxonomy node.
+We create our gazetteer tree based on Geonames\footnote{http://download.geonames.org/export/dump/} and Wikipedia\footnote{http://www.wikipedia.org/}.
 However, in \cite{yin2014pinpointing} the gazetteer hierarchy presents four levels where the lowest level represents a specific point of interest.
-In our approach, we use a subset of the gazetteer hierarchy with three levels, in which the lowest level is represented by a city since a large amount of users specify their location down to this level \cite{hecht2011tweets}.
+In our approach, we use a subset of the gazetteer hierarchy with only three levels: {\em city}, {\em state} and {\em country}.
+We do so because a large amount of users specify their location down to city level \cite{hecht2011tweets}.
 For example, if we have the \textit{city:Manchester}, we associate this location with \textit{region-state:North West} and also with \textit{country:England}.
-Furthermore, the name of locations are considered just in the native language of the country.
+As indicated in our data pre-processing stage, we consider only locations in the native language of the country.
 For instance, in the case of Italy locations, we consider \textit{Roma} and not \textit{Rome}.
 
-\subsubsection{Finding Locations in Microblog Metadata}
+\subsubsection{Location Extraction}
 
-The structure of the Tweet metadata obtains information about the message and the user who shares a message.
-Considering the geographical hierarchy explained in the previous section, we search these locations on different levels of metadata and create one signal for each as following:
+The structure of the tweet metadata contains information about the message and the user.
+Given a small portion of users sharing their current location using GPS coordinates \cite{graham2014world}, we do not consider this level of the tweet metadata in this work.
+
+Considering the aforementioned geographical hierarchy, we extract locations from different parts of the metadata, creating 3 signals for each location:
 
 \begin{itemize}
-	\item Tweet Text: location is mentioned in the attribute \texttt{text} of tweet object, that is, on the body of the message.
-	\item User Location: location is mentioned in the attribute \texttt{location} inside the \texttt{user object}, that is, the location set by the user in his profile.
-	\item Tweet Text - User Location: location is mentioned in the attribute \texttt{text} of tweet object and also location is mentioned in the attribute \texttt{location} inside the \texttt{user object}.
+	\item {\it tweet text:} the location is mentioned in the attribute \texttt{text} of the tweet, that is, on the body of the message.
+	\item {\it user location:} the location is mentioned in the attribute \texttt{location} inside the \texttt{user object}, that is, the location set by the user in their profile.
+	\item {\it tweet text - user location:} the location is mentioned in the attribute \texttt{text} of tweet object and also location is mentioned in the attribute \texttt{location} inside the \texttt{user object}.
 This means that the location is mentioned in the body of the message and the user who shares message has the same location in his profile.
@@
 
-In this way, mixing geographical hierarchy and locations in microblog metadata, we create $NxM$ signals where $N$ is the number of locations obtained by gazetteer tree and $M$ is the number of metadata-levels extracted from Tweet Object.
-For instance, we create a signal for \textit{city:Manchester} and we find this hierarchy in \textit{metadata:Tweet Text} and also in \textit{metadata:User Location}.
+In this way, mixing geographical hierarchy and locations in microblog metadata, we create $NxM$ signals where $N$ is the number of locations obtained by gazetteer tree and $M$ is the number of metadata-levels extracted from the tweet object.
+For instance, we create a signal for \textit{city:Manchester} and we find this hierarchy in \textit{metadata:Tweet text} and also in \textit{metadata:User location}.
 That means that we track the mention of \textit{city:Manchester} at the level of the body of message and at the level of the location of the user profile individually.
 
-\subsection{Time-Windows}
+\subsection{Time-Window}
 In this stage we address the problems of how to divide and determine the time-window size to detect a new emergency situation and what features by the time-window allow it.
@@
 
-According to \citeauthor{guzman2013line} \cite{guzman2013line}: ``If the window size is too small, the ocurrence of empty windows for a term increases, making the noise rate increase and frequency rate tend towards zero.
+According to \citeauthor{guzman2013line} \cite{guzman2013line}: ``If the window size is too small, the occurrence of empty windows for a term increases, making the noise rate increase and frequency rate tend towards zero.
 On the other hand, if the window size is too large, the stability of the signal becomes constant and bursty keyword detection is delayed''.
@@
 
-To characterize the urgency of the messages during a time-window, we use our previous results \cite{DBLP:conf/fdia/Sarmiento17} to compute the \textit{interarrival time} which is defined as $d_{i} = t_{i+1} - t_{i}$, where $d_{i}$ denotes the difference between two consecutive social media messages $i$ and $i+1$ that arrived in moments $t_{i}$ and $t_{i+1}$ respectively.
-Using this definition and according the work of \citeauthor{kalyanam2016prediction} \cite{kalyanam2016prediction}, high-activity events have a high-frequency in the first bins represented by values $d_{i} \approx 0$.
+To characterize the urgency of the messages during a time-window, we compute the \textit{interarrival time} which is defined as $d_{i} = t_{i+1} - t_{i}$, where $d_{i}$ denotes the difference between two consecutive social media messages $i$ and $i+1$ that arrived in moments $t_{i}$ and $t_{i+1}$ respectively.
+Using this definition, which follows the work of \citeauthor{kalyanam2016prediction} \cite{kalyanam2016prediction}, high-activity events have a high-frequency in the first bins represented by values $d_{i} \approx 0$.
 
@@
 
-%\subsection{Subsetting Signals}
-%\begin{itemize}
-%	\item Several signals have a null values along the time.
-For example, any city that is not a big city or capital city, does not a continuous frequency.
-To avoid this behaviour, we use cities just for compute the state frequency signal and later we discard them.
-Finally, we just monitor states and country hierarchy.
-%	\item Furthermore, we remove gps signal because just 1-3 percent of messages have gps activated.
-%\end{itemize}
-
 \subsection{Geographic Spread}\label{sssec:geospread}
@@
 Using this definition, we extend this concept to represent neighborhoods between locations obtained from section \ref{sssec:geohie}.
-For that purpose, we create an \textit{adjancency matrix} $M$, where $M_{i,j} = 1$ represents if two locations are geographically connected and $M_{i,j} = 0$ if they are not connected.
-For instance, if an event is diffused (e.g., earthquake), the detection should be in adjacent-locations independentely of metadata-level.
+For that purpose, we create an \textit{adjacency matrix} $M$, where $M_{i,j} = 1$ represents if two locations are geographically connected and $M_{i,j} = 0$ if they are not connected.
+For instance, if an event is diffused (e.g., earthquake), the detection should be in adjacent-locations independently of metadata-level.
 On the other hand, if an event is focalized (e.g., terrorist attack), just one location should be detected but in different metadata-levels simultaneously.
+
+\section{Experimental Analysis}
+
+Our experiments aim to find empirical evidences that allow detecting an emergency situation.
+Thereby, we use the locations for a specific country in different levels of the geographic hierarchy and the tweet metadata.
+We construct our ground truth based on two publicly available earthquake catalogs and apply the proposal presented in the above section.
+Thereafter, we study the performance of our methodology mainly based on the concept of affected adjacent locations during an emergency situations.
+
+\subsection{Dataset Description}
+
+We collect data from Twitter Public Streaming API, which allows access to subsets equal to $1\%$ of public status descriptions in real-time.
+With this tool, we can retrieve either messages using a set of keywords or messages from specific locations setting a bounding box.
+In our approach, we get entire subsets of messages without use keywords or specific locations.
+Then, we retrieve random messages about any topic and any place in the world.
 
@@
 
-\section{Dataset}
-
-Data is collected from Twitter Public Streaming API, which allows access to subsets equal to $1\%$ of public status descriptions in real-time.
-With this tool, we can retrieve either messages using a set of keywords or messages from specific locations setting a bounding box.
-In our approach, we get entire subsets of messages without use keywords or specific locations.
-Then, we retrieve random messages about any topic and any place in the world.
-
 \subsection{Ground Truth}
 
-We analyse five earthquakes with magnitudes between $5.5Mw$ and $7.6Mw$\footnote{Mw: the moment magnitude scale}, ocurring in Italian-speaking and Spanish-speaking countries between October 2016 and April 2017 (Table \ref{tab:eqs}).
+We analyze five earthquakes with magnitudes between $5.5Mw$ and $7.6Mw$\footnote{Mw: the moment magnitude scale}, occurring in Italian-speaking and Spanish-speaking countries between October 2016 and April 2017 (Table \ref{tab:eqs}).
 For that purpose, we collect $20$ million of messages 12-hours before and after the emergency situation events.
@@
 
-\section{Experiments}
+\subsection{Methodology}\label{sec:methodology}
+
+According to the ``data classification'' module (Figure \ref{fig:diagram}), we first train a classifier to identify emergency events.
+Also, we introduce the hierarchy dependence to understand the local and national impact when an high-impact real-world event occurs.
+Besides, diffused and focalized events are identified with the goal of filtering false positives detections.
 
@@
 To classify messages we employed traditional binary classifier Support Vector Machine \textit{(SVM)}.
-As a result of the analysed data scattering (Figure \ref{fig:scatter}), we separated country and state in different datasets and set both kernels and classification parameters independentely.
+As a result of the analyzed data scattering (Figure \ref{fig:scatter}), we separated country and state in different datasets and set both kernels and classification parameters independently.
 On the one hand, \textit{country} classifier uses a polynomial kernel and strict-parameters for gamma, cost and weights since a great amount of messages are included in country hierarchy as an effect of the minor hierarchies.
@@
 Given that an emergency situation is not an usual event, we have an highly unbalanced data respect to the classes after labeled ($1\approx2\%$ of positive class corresponding to \textit{detection}).
-Therefore, we used \textit{under-sampling} \cite{lunardon2014rose} over country and state datasets increasing our possitive class to $15\approx18\%$.
+Therefore, we used \textit{under-sampling} \cite{lunardon2014rose} over country and state datasets increasing our positive class to $15\approx18\%$.
 Additionally, to validate our model, we used \textit{5-fold cross-validation} where one earthquake dataset is used as testing and the remaining earthquakes dataset as training.
@@
 For this analysis, we search the time-windows for each hierarchy where the all metadata-levels are well classified with correct class.
-According to the results shown on the middle of the Table \ref{tab:perfomance}, when we analyse country and state independentely the values of Precision, F1 and FPR have worst values than the analysis by label .
+According to the results shown on the middle of the Table \ref{tab:perfomance}, when we analyze country and state independently the values of Precision, F1 and FPR have worst values than the analysis by label .
 \begin{table}
@@
 In contrast to the the previous analyses, we improved the values of the Precision, F1 and FPR.
-The last metrics are very important because there are no time-windows incorrectly assigned as emergency situations.
+The last metric is very important because there are no time-windows incorrectly assigned as emergency situations.
 Consequently, the Recall values decrease which means that our method remove some time-windows classified as detection.
@@
 Similar to  \textit{Country(2)-State with G.S.}, we improved the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with a average delay equal to $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the first detection.
+
+\section{Results}
+
+In this section we present an evaluation of our methodology for emergency situation detection.
+We evaluate our method using different events occurred in England related to terrorist attacks and non emergency situations.
+In this way, our main goal is to detect emergency events in non labeled data and reduce (as much as possible) the false positives detections.
+For these tasks, we use the analysis with the best performance presented in the Section \ref{sec:methodology}.
 
@@
 As can be noted on Table \ref{tab:online1} and Table \ref{tab:online2}, we study three terrorist attacks and five high-impact real-world events related to soccer matches, music concerts and political elections.
-In the case of Premier League Soccer Matches and U.K Elections, we can not identify the beginning of the event, since in the first one there are many soccer matches during the analysed day and in the second one there is no a specific start time.
+In the case of \textit{Premier League Soccer Matches} and \textit{U.K Elections}, we can not identify the beginning of the event, since in the first one there are many soccer matches during the analyzed day and in the second one there is no a specific start time.
 In order to know the topics when our method detects an event, we computed the Top 3 Bigrams in the detected time-windows.
@@
 These detections are related to the event given that the bigrams represent terms associated with crisis situations.
-However, the London Terrorist Attack has $50\%$ of the detected time-windows after the event, which means that there are seven time-windows non-related to emergency situations.
+However, the \textit{London Terrorist Attack} has $50\%$ of the detected time-windows after the event, which means that there are seven time-windows non-related to emergency situations.
 Besides the crisis situations analysis, we also study the number of detected time-windows in non-related to emergency situation events.
@@
 On the other hand, the second evaluation \textit{Country(3)-State with G.S.} decreases the number non-related to emergency situations events detected as crisis situations.
-We can see three time-windows in two events detected as emergency situations (England vs Eslovenia, and Metallica Live in London).
+We can see three time-windows in two events detected as emergency situations (\textit{England vs Slovenia}, and \textit{Metallica Live in London}).
 In these cases, the time-windows are detected before the event and corresponding a non emergency situations according to the bigrams.
-Furthermore, when we analysed the number of the detected emergency situations, two-thirds ($66\%$) of the events are detected correctly with average delay time equal to $30.3$ minutes.
-In the case of London Terrorist Attack, our method detects one time-window before the event but the bigrams describe that the detections do not correspond to crisis situations.
+Furthermore, when we analyzed the number of the detected emergency situations, two-thirds ($66\%$) of the events are detected correctly with average delay time equal to $30.3$ minutes.
+In the case of \textit{London Terrorist Attack}, our method detects one time-window before the event but the bigrams describe that the detections do not correspond to crisis situations.
 
@@
 Also, our method is independent of the textual features because we apply the model over different languages as Spanish, Italian and English.
-Furthermore, we test our model in different types of events such as earthquakes and terrorist attacks, and also on different magnitudes (in the case of earthquakes) and number of affected people (e.g., Manchester Terrorist Attack vs Westminster Terrorist Attack).
+Furthermore, we test our model in different types of events such as earthquakes (EQ) and terrorist attacks (TA), and also on different magnitudes (in the case of earthquakes) and number of affected people (e.g., \textit{Manchester Terrorist Attack} vs \textit{Westminster Terrorist Attack}).
 
@@
 
-%\begin{itemize}
-%	\item explain that we dont need compute textual features as bigrams, top-k terms.
-%	\item explain that we just need de gazetteer tree
-%	\item explain when our methodology could be not working. eg. why not working correctly in england. country with many language. How event like soccer match are filtered and when not working
-%	\item Delay time: explain difference between focalized and diffused event. People feel the event at the same time, activity in affected locations.
-%	\item Number of locations affected in the first and second detection for focalized and diffused (plot)
-%	\item Diffused: mercalli and locations detected.
-
-%\end{itemize}
-
 \section{Conclusion}

### 2018-04-08 paper.tex
Add example of gazetteer tree, E.S definition and explanation about baseline

@@
 \usepackage{array,multirow}
+\usepackage{todonotes}
 % Copyright
@@
 % DOI
-\acmDOI{10.475/123_4}
+%\acmDOI{10.475/123_4}
 
 % ISBN
-\acmISBN{123-4567-24-567/08/06}
+%\acmISBN{123-4567-24-567/08/06}
 
@@
 \affiliation{%
-  \institution{Department of Computer Science, University of Chile}
+  \institution{Department of Computer Science \\ University of Chile}
   \streetaddress{P.O. Box 1212}
@@
 \affiliation{%
-  \institution{Department of Geophysics, University of Chile}
+  \institution{Department of Geophysics \\ University of Chile}
 %  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
@@
 
-\keywords{Emergency Situations, Citizen Sensors, Social Media}
+\keywords{Emergency Situations; Citizen Sensors; Social Media}
+
+\copyrightyear{2018} 
+\acmYear{2018} 
+\setcopyright{acmlicensed}
+\acmConference[WebSci'18]{10th ACM Conference on Web Science}{May 27--30, 2018}{Amsterdam, Netherlands}
+\acmBooktitle{WebSci'18: 10th ACM Conference on Web Science, May 27--30, 2018, Amsterdam, Netherlands}
+\acmPrice{15.00}
+\acmDOI{10.1145/3201064.3201077}
+\acmISBN{978-1-4503-5563-6/18/05}
 
@@
 
-During emergency situations, traditional media may suffer infrastructure issues and real-time communications could be disrupted.
+\todo{definicion}  \textcolor{red}{According to United Nations Department of Humanitarian Affairs \cite{dha1992internationally}, an emergency situation can be defined as ``a sudden and usually unforeseen event that calls for immediate measures to minimize its adverse consequences''}.
+During these events, traditional media may suffer infrastructure issues and real-time communications could be disrupted.
 Instead, microblogging has played a critical role over the last fifteen years allowing users to share real-time information from people local to the incident, such as status updates, casualties, damages and alerts \cite{kumar2011tweettracker,imran2013extracting,stowe2016identifying,reuterfifteen}.
@@
 Our main contribution is to create a methodology that detects instantaneous emergency situations just by using the location related information for a specific country.
-Furthermore, our approach does not depend on language to detect a new event since it does not use textual features as input.
+\textcolor{red}{Since an emergency situation is a high-impact real-world event, we use locations in different levels of geographic hierarchies for detecting anomalies in the lowest levels (as cities or states) and the highest levels (a country) at the same time} \todo{uso de localidades}.Furthermore, our approach does not depend on language to detect a new event since it does not use textual features as input.
 In addition, we characterize focalized and diffused crisis situations.
+
+\todo{sobre el baseline}\textcolor{red}{To validate our hypothesis, we create a ground truth based on random messages retrieving from Twitter Public Stream without specific keywords and geolocations.
+In contrast to previous works where they retrieve messages for a specific domain and type of crisis events, we get domain-independent messages to find and demonstrate our hypothesis that an anomaly in the activity related to locations can detect an emergency situation.
+Hence, we do not compare our detection method with others because they are centered in tracking specific crisis events.
+Besides, there are no jobs with independent domain to use as baseline related to emergency situations.}
 
@@
 As indicated in our data pre-processing stage, we consider only locations in the native language of the country.
-For instance, in the case of Italy locations, we consider \textit{Roma} and not \textit{Rome}.
+For instance, in the case of Italy locations, we consider \textit{Roma} and not \textit{Rome} (Figure \ref{fig:gazetteer}).
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{img/gazetteer_example_paper2.png}
+	\caption{Example of gazetteer tree for Italy.}
+	\label{fig:gazetteer}
+\end{figure}
 
@@
 With this tool, we can retrieve either messages using a set of keywords or messages from specific locations setting a bounding box.
-In our approach, we get entire subsets of messages without use keywords or specific locations.
+In our approach, we get entire subsets of messages without use of keywords or specific locations.
 Then, we retrieve random messages about any topic and any place in the world.
@@
 
-We analyze five earthquakes with magnitudes between $5.5Mw$ and $7.6Mw$\footnote{Mw: the moment magnitude scale}, occurring in Italian-speaking and Spanish-speaking countries between October 2016 and April 2017 (Table \ref{tab:eqs}).
-For that purpose, we collect $20$ million of messages 12-hours before and after the emergency situation events.
+\textcolor{red}{We choose earthquakes as crisis situation for our ground truth since that according to \citeauthor{carr1932disaster} \cite{carr1932disaster}, these events are also known as \textit{instantaneous-diffused} events.
+These events are characterized since that nobody could do anything about it and wrecked its effects on the entire community.
+This definition is relevant because an unexpected (or instantaneous) event generates an anomaly of frequency in the social media activity since that it disrupts the ``normal life'' of the users.
+Furthermore, diffused events affect a big portion of the users simultaneously generating a collective reaction in neighbors affected locations.}
 
-According our proposal we create both the gazetteer hierarchies\footnote{http://users.dcc.uchile.cl/$\sim$hsarmien/gazetteer.html} for each country and construct the signals based on each hierarchy and metadata-level.
+In this way, we analyze five earthquakes with magnitudes between $5.5Mw$ and $7.6Mw$\footnote{Mw: the moment magnitude scale}, occurring in Italian-speaking and Spanish-speaking countries between October 2016 and April 2017 (Table \ref{tab:eqs}).
+For that purpose, we collect $20$ million messages 12 hours before and after the emergency situation events.
+
+According to our proposal we create both the gazetteer hierarchies\footnote{http://users.dcc.uchile.cl/$\sim$hsarmien/gazetteer.html} for each country and construct the signals based on each hierarchy and metadata-level.
 As a result of the number of messages of each signal (Table \ref{tab:message_signal}), we discard all signals related to city hierarchy since a great amount of small cities have zero frequency in a normal situation unlike to capital or metropolitan cities.
@@
 
-The exactly datetime event is obtained from National Seismology Agency in Chile\footnote{http://www.sismologia.cl/} and 
+The exact event datetime is obtained from National Seismology Agency in Chile\footnote{http://www.sismologia.cl/} and 
 National Institute of Geophysics and Volcanology in Italy\footnote{http://www.ingv.it/it/}.
@@
 	\includegraphics[width=\columnwidth]{img/labeled2.png}
-	\caption{Average variation in emergency situations between time-windows.}
+	\caption{Average variation in emergency situations between time-windows.
+\textcolor{red}{Positive and negatives values in x-axis represent the following and previous time-windows from the beginning of the event respectively.}}
 	\label{fig:labeled}
@@
 
-Given that an emergency situation is not an usual event, we have an highly unbalanced data respect to the classes after labeled ($1\approx2\%$ of positive class corresponding to \textit{detection}).
+Given that an emergency situation is not a usual event, we have an highly unbalanced data respect to the classes after labeled ($1\approx2\%$ of positive class corresponding to \textit{detection}).
 Therefore, we used \textit{under-sampling} \cite{lunardon2014rose} over country and state datasets increasing our positive class to $15\approx18\%$.
@@
 \begin{table*}
-	\caption{Online evaluation by time-windows (T-W) using Country(2)-State with G.S. method}
+	\caption{On-line evaluation of events occurred in England by time-windows (T-W) using Country(2)-State with G.S.
+method.
+The table shows the total number of detected time-windows, the number of detected time-windows before to the beginning and after to the end of the event.
+Last two columns show the detection delay time respect to the beginning of the event and the top 3 bigrams when the detection occurs.}
 	\label{tab:online1}
@@
 \begin{table*}
-	\caption{Online evaluation by time-windows (T-W) using Country(3)-State with G.S. method}
+	\caption{On-line evaluation of events occurred in England by time-windows (T-W) using Country(3)-State with G.S.
+method.
+The table shows the total number of detected time-windows, the number of detected time-windows before to the beginning and after to the end of the event.
+Last two columns show the detection delay time respect to the beginning of the event and the top 3 bigrams when the detection occurs.}
 	\label{tab:online2}
@@
 
-Considering the geographic spread by states and the number of metadata-levels by country hierarchy, we analysed the results shown on the bottom of the Table \ref{tab:perfomance}.
+Considering the geographic spread by states and the number of metadata-levels by country hierarchy, we analyzed the results shown on the bottom of the Table \ref{tab:perfomance}.
 On the one hand, the row with value equal to \textit{Country(2)-State with G.S.} represents the detection when we considered at least two metadata-levels for the country hierarchy and the geographic spread for states.
@@
 Consequently, the Recall values decrease which means that our method remove some time-windows classified as detection.
-Beside the percent of emergency situations detected is equal to $100\%$ with a average delay equal to $10.4$ minutes ($min=6$, $max=14$) from the impact of the event to the first detection.
+Beside the percent of emergency situations detected is equal to $100\%$ with an average delay equal to $10.4$ minutes ($min=6$, $max=14$) from the impact of the event to the first detection.
 
 On the other hand, \textit{Country(3)-State with G.S.} represent the detection when we considered three metadata-levels for country hierarchy and the geographic spread for states.
-Similar to  \textit{Country(2)-State with G.S.}, we improved the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with a average delay equal to $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the first detection.
+Similar to  \textit{Country(2)-State with G.S.}, we improved the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with an average delay equal to $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the first detection.
 
@@
 
-In this section we present an evaluation of our methodology for emergency situation detection.
+In this section, we present an on-line evaluation of our methodology for emergency situation detection.
+We identify several events occurred in England that includes emergency situations and non-emergency situations.
+In this way, our main goal is to detect emergency events in non labeled data and using a different language with respect to the languages used in our classifier.
+To reduce (as much as possible) the false positive detections, we use the analysis with the best performance presented in the Section \ref{sec:methodology}.
+
+%o reduce (as much as possible) the false positives detections. For these tasks, we use the analysis with the best performance presented in the Section \ref{sec:methodology}.
+
+%In this section we present an evaluation of our methodology for emergency situation detection.
 We evaluate our method using different events occurred in England related to terrorist attacks and non emergency situations.
@@
 
-\subsection{Online Evaluation}
-For our evaluation in the Twitter Public Stream, we trained classifier with five earthquakes identified in our ground truth.
-Furthermore, our evaluation dataset is formed by eight different events that occurred in England between December 2016 and October 2017.
-For each event we considered the full-day in which they ocurred.
+\subsection{On-line Evaluation}
+For our evaluation in the Twitter Public Stream, we trained a classifier with five earthquakes identified in our ground truth.
+Furthermore, our on-line evaluation dataset is formed by eight different events that occurred in England between December 2016 and October 2017.
+For each event we considered the full-day in which they occurred.
 The main goals of this evaluation is to know the capacity of our method to detect emergency situations and discard those non-related to emergency events that involve location references.
@@
 As can be noted on Table \ref{tab:online1} and Table \ref{tab:online2}, we study three terrorist attacks and five high-impact real-world events related to soccer matches, music concerts and political elections.
-In the case of \textit{Premier League Soccer Matches} and \textit{U.K Elections}, we can not identify the beginning of the event, since in the first one there are many soccer matches during the analyzed day and in the second one there is no a specific start time.
+\textcolor{red}{In the case of the terrorist attacks, we study this type of events since that according to \citeauthor{carr1932disaster} \cite{carr1932disaster}, these crisis are identified as \textit{instantaneous-focalized} events, where an unexpected event affects to the community but in a reduced area.
+Unlike to earthquakes (identified as \textit{instantaneous-diffused} events), we study the capacity of our classifier to detect another type of event where the number of affected people are smaller than earthquakes, tsunami, and others \textit{instantaneous-focalized} events}.
+
+For \textit{Premier League Soccer Matches} and \textit{U.K Elections}, we can not identify the beginning of the event, since in the first one there are many soccer matches during the analyzed day and in the second one there is no a specific start time.
 In order to know the topics when our method detects an event, we computed the Top 3 Bigrams in the detected time-windows.
-Also, we calculated the delay time just for emergency events.
+Also, we calculate the delay time for emergency events since the beginning of the event until the first detection.
 
@@
 Also, our method is independent of the textual features because we apply the model over different languages as Spanish, Italian and English.
-Furthermore, we test our model in different types of events such as earthquakes (EQ) and terrorist attacks (TA), and also on different magnitudes (in the case of earthquakes) and number of affected people (e.g., \textit{Manchester Terrorist Attack} vs \textit{Westminster Terrorist Attack}).
+Furthermore, we test our model in different types of crisis events such as earthquakes (EQ) and terrorist attacks (TA), \textcolor{red}{ where these are identified in the literature as \textit{instantaneous-diffused} and \textit{instantaneous-focalized} events respectively}.
+Also, we apply our methodology on different magnitudes (in the case of earthquakes) and number of affected people (e.g., \textit{Manchester Terrorist Attack} vs \textit{Westminster Terrorist Attack}).
 
@@
 
-%\begin{acks}
-
-%\end{acks}
+\begin{acks}
+This work has been partially funded by the Millennium Institute for Foundational Research on Data and FONDECYT 1170218.	
+\end{acks}
 

### 2019-10-01 paper.tex
first commit

--- 2018-02-05 paper.tex
@@
 \usepackage{array,multirow}
+\usepackage{balance}
 % Copyright
@@
 % DOI
-\acmDOI{10.475/123_4}
+%\acmDOI{10.475/123_4}
 
 % ISBN
-\acmISBN{123-4567-24-567/08/06}
-
-%Conference
-%\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El Paso, Texas USA} 
-%\acmYear{1997}
-%\copyrightyear{2016}
-
-%\acmArticle{4}
-%\acmPrice{15.00}
-
-% These commands are optional
-%\acmBooktitle{Transactions of the ACM Woodstock conference}
-%\editor{Jennifer B. Sartor}
-%\editor{Theo D'Hondt}
-%\editor{Wolfgang De Meuter}
 
 \begin{document}
-%\title{Detecting Emergency Situations Based on Locations in Twitter}
-%\title{Detecting Emergency Situations in Twitter Using Self-Organized Citizen Sensors}
-\title{Domain-Independent Detection of Emergency Situations\\Based on Social Activity Related to Geolocations}
-
-%\titlenote{Produces the permission block, and
- % copyright information}
-
-%\subtitle{Extended Abstract}
-%\subtitlenote{The full version of the author's guide is available as
-  %\texttt{acmart.pdf} document}
+\title{A Cross-Domain Exploratory Analysis of \\ Social Media Data During Crisis Events}
 
 \author{Hernan Sarmiento}
-%\authornote{Dr.~Trovato insisted his name be first.}
-%\orcid{1234-5678-9012}
 \affiliation{%
-  \institution{Department of Computer Science, University of Chile}
-  \streetaddress{P.O. Box 1212}
+  \institution{Department of Computer Science \\ University of Chile}
+  %\streetaddress{P.O. Box 1212}
    \city{Santiago} 
@@
 \author{Barbara Poblete}
-%\authornote{The secretary disavows any knowledge of this author's actions.}
 \affiliation{%
-  \institution{Department of Computer Science, University of Chile}
-  \streetaddress{P.O. Box 1212}
+  \institution{Department of Computer Science \\ University of Chile}
+  %\streetaddress{P.O. Box 1212}
    \city{Santiago} 
@@
 
-\author{Sergio Campos}
-%\authornote{This author is the
- % one who did all the really hard work.}
-\affiliation{%
-  \institution{Department of Geophysics, University of Chile}
-%  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Santiago} 
-  \country{Chile}}
-\email{jaime@dgf.uchile.cl}
+\begin{abstract}
 
-% The default list of authors is too long for headers.
-%\renewcommand{\shortauthors}{B. Trovato et al.}
-
-\begin{abstract}
-%Most methods to detect emergency situations using Twitter rely on keyword.
-The problem of keyword-based methods is the need to train in specific domains and language for different type of events, for example: earthquakes, typhoons, terrorist attacks, tornadoes, etc.
-Our proposal is to rather use the recurring mention of a country locations in microblogging messages to identify such events without using keywords and characterize through of interarrival time probability distributions.
-Our method use a SVM classifier that is independent of the text language of the messages and can be training and testing among different languages and domains.
-%
-Most methods for detecting emergency situations using Twitter rely on identifying features within messages that contain domain-specific keywords.
-However, keyword-based methods require models to be trained on historical data of specific domains, in multiple languages, and for different types of events (e.g., earthquakes, floods, wildfires, etc.).
-%
-In addition to being costly, these approaches may fail to detect previously unexpected situations, such as uncommon catastrophes or terrorist attacks.
-%
-Nevertheless, collective mentions of keywords are not the only type of self-organizing phenomena that may arise when a real-world extreme situation occurs.
-%
-Just as physical sensors become activated when stimulated, localized citizen sensors will also react in a similar manner.  
-%
-To leverage this information, we propose to use self-organized geolocation related activity to identify emergency situations.
-%
-We propose to detect such events by tracking the frequencies, and probability distributions of the interarrival time of the messages related to specific locations.
-%
-Using an off-the-shelf classifier that is independent of domain-specific features, we study and describe emergency situations based solely on location-based features in messages.
-%
-Our findings indicate that anomalies in location-related social media user activity indeed provide information for automatically detecting emergency situations independent of their domain.
 \end{abstract}
@@
 
-\keywords{Emergency Situations, Citizen Sensors, Social Media}
+\keywords{Emergency Situations; Citizen Sensors; Social Media}
 
@@
 \section{Introduction}
-Social media has become a major communication channel during high-impact real-world events, such as elections, sports events, emergency situations, among others.
-In each case, users act as ``citizen sensors'' sharing and posting their mood, opinion, photos, videos and geographical points of interest.
 
-During emergency situations, traditional media may suffer infrastructure issues and real-time communications could be disrupted.
-Instead, microblogging has played a critical role over the last fifteen years allowing users to share real-time information from people local to the incident, such as status updates, casualties, damages and alerts \cite{kumar2011tweettracker,imran2013extracting,stowe2016identifying,reuterfifteen}.
-For this reason, researchers have studied user behavior during these events to detect, summarise and classify messages with the goal of helping authorities, and the general public, with situational awareness to provide fast and conscientious responses during crisis situations.
-
-Twitter is a microblogging platform that allows users to share short messages (called \textit{tweets}) and is currently used worldwide by over $300$ million people.\footnote{http://www.statista.com/statistics/282087/number-of-monthly-active-twitter-users/} About $80\%$ of Twitter users access from mobile devices, which contributes to the immediacy of diffusion of information, especially during crisis situations \cite{castillo2016big}.
-One of the most important tasks during these situations, is to timely detect incoming real-world events.
-This is because, as with most social media conversations, messages are often overridden with irrelevant and redundant noise.
-In current works \cite{kumar2011tweettracker,caragea2011classifying,imranaidr2014,maldonado2017}, these tasks are solved with methods that rely on keyword based filters on the \textit{Twitter Public Streaming API}.\footnote{https://developer.twitter.com/en/docs/tweets/filter-realtime/overview} The problem with keyword-based methods is the need to train in specific domains for different type of events.
-For instance, \citeauthor{olteanu2014} \cite{olteanu2014} generate a set of keywords based on different datasets.
-However, this is not sufficient for cases in which specific and previously unseen terms arise for a particular event (e.g., \textit{\#eqnz} for Earthquakes in New Zealand, or \textit{\#pabloph} for Typhoon Pablo in Philippines) \cite{potts2011tweeting,bruns2012local,karimi2013classifying}.
-Furthermore, these sets of keywords are commonly obtained a particular language and will not work in others.
-
-Additionally, several works related to crisis situations show a strong relationship between a type of event (what) and the spatiotemporal dimensions (when and where).
-For example, it has been observed that top trends during earthquakes are related to location mentions \cite{mendoza2010twitter}, also, that there is a strong relationship between the proximity to a hurricane path and hurricane-related social media activity \cite{kryvasheyeu2016rapid}.
-Other works have addressed the extraction of locations and points of interest during floods \cite{lingad2013location}, as well as mixing geographic information system (GIS) information with geo-tagged messages to improve disaster mapping and real-time event tracking \cite{huang2015disastermapper}.
-
-Based on previous works, the question we address here is: \textit{Is there evidence to detect an emergency situation based on anomalies in messages related to contain locations?}.
-
-Here, we propose a method based on recurring references to country level locations in message metadata.
-For this task, we create a gazetteer tree based on the hierarchy of the specific country, divide the messages into fixed time-windows and compute the frequency and the probability distributions of the interarrival time of the messages for each geographical hierarchy.
-To detect an emergency situation, we train a SVM classifier by each hierarchy and apply a geographic spread to filter false positives detection considering that an emergency situation can be \textit{Focalized} or \textit{Diffused}.
-
-Our main contribution is to create a methodology that detects instantaneous emergency situations just by using the location related information for a specific country.
-Furthermore, our classifiers do not depend on language to detect a new event since they do not use textual features as input.
-Moreover, we characterize crisis situations that affect small or large geographical areas.
-
-The paper is organized as follows: we first introduce an overview of relevant literature related to emergency situation and event detection in social media.
-Next we present a complete description of our proposal divide in four parts.
-Thereafter, we provide a full description of our dataset and later we summarise our experimental validation over the ground truth and on-line evaluation.
-Finally, we deliver our discussions, conclusions and future work.
-
-\section{Related Work}
-
-We discuss relevant work for our proposal, primary from two areas: (1) crisis-related social media monitoring and (2) event detection based on locations.
-
-\subsection{Crisis-Related Social Media Monitoring}
-
-Twitter has been used extensively during emergency situations to extract and identify relevant information.
-However, social media communications during emergency situations have become so abundant that it is necessary to sift through millions of data points to find information that is most useful during a given event \cite{imran2015processing}.
-
-One of the main tasks related to emergency situations is to detect a new real-crisis event in social media.
-Most existing event-detection methods described in the literature are based on keywords.
-For instance, \textit{TweetTracker} \cite{kumar2011tweettracker} presented a case study using tweets that discuss a cholera outbreak in Haiti.
-The primary mechanism for monitoring tweets is through specific keywords and hashtags filters related to the Haiti.
-To detect a new event, emerging trends are identified based on the analysis of older tweets.
-In the same way, \textit{EMERSE} \cite{caragea2011classifying} used a set of keywords related to the Haiti earthquake and applies a random forests algorithm for classifying messages.
-Likewise, the \textit{Twicalli} system \cite{maldonado2017} introduced an unsupervised approach to detect earthquakes that only requires a general list of keywords.
-
-Researchers at CSIRO Australia proposed \textit{ESA} \cite{cameron2012emergency,yin2012esa}, a system to detect disasters in Australia and New Zealand.
-This system is based on a probabilistic method to identify bursty keywords and historical data to build a language model of word occurrences.
-Alerts are identified if a term has a probability distribution that significantly deviates from the language model.
-Similarly, \textit{Twitcident} \cite{abel2012twitcident} system was developed for detecting incidents which relies on emergency broadcasting services, such as the police, the fire department and other public emergency services.
-The \textit{Twitcident} framework translates the broadcasted message into an initial incident profile that is applied as query to collect messages from Twitter, where an incident profile is a set of weighted attribute-value pairs that describe the characteristics related to the incident.
-
-\textit{CrisisLex} \cite{olteanu2014} is introduced to extract different crisis-related terms to filter messages in Twitter. In this work,
-authors collected six disasters which affected several millions of people.
-Data was collected from Twitter using two samples: a keyword-based sample and location-based sample.
-Based on datasets, they created a lexicon of the most frequent terms that appear in relevant messages posted during different types of crisis situations.
-
-Our current work, on the other hand, is based on detecting events by monitoring \textit{zscore} variations through fixed time-windows.
-
-\subsection{Event Detection Based on Locations}
-
-In addition to crisis-related social media monitoring, there are also some unsupervised event detection approaches based on location information.
-\textit{Jasmine} \cite{watanabe2011jasmine} detected local real-world events using geolocation information from microblog documents.
-To detect such events, they identify a group of tweets that describe a particular theme, which are generated within a short time frame and a same geographic area.
-Similarly, \citeauthor*{unankard2015emerging} \cite{unankard2015emerging} proposed an approach for early detection of emerging hotspot events in social networks with location sensivity.
-In this work, the authors identified strong correlations between user locations and event locations when detecting emerging events using content similarity between clusters.
-In the work of \citeauthor{walther2013geo} \cite{walther2013geo} real-world events were detected in a small scale with messages from the New York metropolitan area.
-There, clusters were created for each candidate event and evaluated using cluster score based on textual features (sentiment analysis, common theme, duplicate, etc) and other features (tweet count, unique coordinates, etc.).
-
-Supervised approaches have also been proposed.
-\textit{TEDAS} \cite{li2012tedas} detected, analysed and identified events using refined rules (e.g., keywords, hashtags) and classify messages based on content as well Twitter specific features as URLs, hashtags and mentions.
-Besides, location information is extracted using both explicit geographical coordinates and implicit geographical references in the content.
-In the same way, \citeauthor{becker2011beyond} \cite{becker2011beyond} proposed an on-line clustering technique, which continuously clusters similar tweets and then classifies the clusters using Support Vector Machine algorithm.
-Finally, events (clusters) are classified into real-world events or nonevents.
-
-\medskip
-
-\noindent{\textbf{Summary.}} Most previous works to detect an emergency situations rely on keywords using probabilistic temporal models for specific domains (e.g., keywords or location).
-In general terms, these approaches need knowledge about the event and a new emerging topic referring to a crisis cannot be identified to detect an emergency situation.
-In the case of the works on event detection based on locations, they are designed to be applied in a small area and based on historical data.
-Regardless of which approach is implemented (supervised or unsupervised), textual features are the most used attributes to characterise an event and on-line clustering is the most common technique to create candidate events.
-
-Based on the works presented by \citeauthor{guzman2013line} \cite{guzman2013line} and \citeauthor{maldonado2017} \cite{maldonado2017}, we extend the ideas of bursty keywords and \textit{z-score} variation between fixed time-windows and we apply these proposals over locations to identify anomalies in social media activity.
-Furthermore, we do not use a set of keywords to detect specific types of events (e.g., earthquake, flood or terrorist attack).
-In contrast, our proposal is focused on detecting such events by tracking frequencies, and probability distributions of the interarrival time of the messages related to specific locations.
-
-\section{Proposed Approach}
-
-In order to provide a complete coverage of location-based detection of emergency situations, we propose an approach which has four stages (depicted in the Data Process module in Figure \ref{fig:diagram}).
-Next, we describe each stage:
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{img/diagram2.png}
-	\caption{Key components of the proposed approach.}
-	\label{fig:diagram}
-\end{figure}
-
-\subsection{Data Pre-Processing}
-%In this stage, data is pre-processed from Twitter Streaming API\footnote{algo}, which allows access to subsets equal to $1\%$ of public status descriptions in real-time.
-With this tool, we can be retrieve messages using a set of keywords or messages from specific locations setting a bounding box.
-In our approach, we get entire subsets of messages without use keywords or specific locations.
-Then, we retrieve random messages about any topic and any place in the world.
-
-%To increase the number of messages related to locations, we consider tweets and retweets messages. 
-Since we consider users as citizen sensors, we filter messages depending of native language of each country using the attribute \textit{lang} retrieved from the metadata in tweets.\footnote{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object} For example, if we analyse Italy, we just consider messages in Italian.
-Finally, we remove user mentions, URLs,  special characters and apply text tokenization.
-Also we do not remove hashtags or stopwords because some locations can be included in hashtags terms and some places can contain stopwords that can differentiate them from other similar non-related to locations terms.
-
-\subsection{Signal Creation}
-The problems that we address in this stage are how to infer locations from microblog messages to track signals over time in the data stream and what is the lowest hierarchy level to find and assign to these signals.
-
-\subsubsection{Geographical Hierarchy}\label{sssec:geohie}
-Using the idea of \textit{Gazetteer as a Tree} presented in \cite{yin2014pinpointing} in which each place is associated with a canonical taxonomy node, we construct our gazetteer tree based on Geonames\footnote{http://download.geonames.org/export/dump/} and Wikipedia\footnote{http://www.wikipedia.org/} for each country to analyse.
-However, in \cite{yin2014pinpointing} the gazetteer hierarchy presents four levels where the lowest level represents a specific point of interest.
-In our approach, we use a subset of the gazetteer hierarchy with three levels, in which the lowest level is represented by a city since a large amount of users specify their location down to this level \cite{hecht2011tweets}.
-For example, if we have the \textit{city:Manchester}, we associate this location with \textit{region-state:North West} and also with \textit{country:England}.
-Furthermore, the name of locations are considered just in the native language of the country.
-For instance, in the case of Italy locations, we consider \textit{Roma} and not \textit{Rome}.
-
-\subsubsection{Finding Locations in Microblog Metadata}
-
-The structure of the Tweet metadata obtains information about the message and the user who shares a message.
-Considering the geographical hierarchy explained in the previous section, we search these locations on different levels of metadata and create one signal for each as following:
+\section{Data Acquisition}
 
 \begin{itemize}
-	\item Tweet Text: location is mentioned in the attribute \texttt{text} of tweet object, that is, on the body of the message.
-	\item User Location: location is mentioned in the attribute \texttt{location} inside the \texttt{user object}, that is, the location set by the user in his profile.
-	\item Tweet Text - User Location: location is mentioned in the attribute \texttt{text} of tweet object and also location is mentioned in the attribute \texttt{location} inside the \texttt{user object}.
-This means that the location is mentioned in the body of the message and the user who shares message has the same location in his profile.
-In this case, tweet text and user location can be different in the smallest hierarchy, but in the highest level can be equal.
+	\item CrisisLex and CrisisNlp.
+	\item XX dataset with YY instances and ZZ events.
 \end{itemize}
 
-In this way, mixing geographical hierarchy and locations in microblog metadata, we create $NxM$ signals where $N$ is the number of locations obtained by gazetteer tree and $M$ is the number of metadata-levels extracted from Tweet Object.
-For instance, we create a signal for \textit{city:Manchester} and we find this hierarchy in \textit{metadata:Tweet Text} and also in \textit{metadata:User Location}.
-That means that we track the mention of \textit{city:Manchester} at the level of the body of message and at the level of the location of the user profile individually.
+\section{Data Preparation}
 
-\subsection{Time-Windows}
-In this stage we address the problems of how to divide and determine the time-window size to detect a new emergency situation and what features by the time-window allow it.
+\begin{itemize}
+	\item Remove duplicates
+	\item Remove data without metadata
+\end{itemize}
 
-\subsubsection{Determining Optimal Window Size}
+\subsection{Detecting Languages}
 
-According to \citeauthor{guzman2013line} \cite{guzman2013line}: ``If the window size is too small, the ocurrence of empty windows for a term increases, making the noise rate increase and frequency rate tend towards zero.
-On the other hand, if the window size is too large, the stability of the signal becomes constant and bursty keyword detection is delayed''.
-Using this definition, we divide our signals into windows of six minutes because it divides a 24-hour day exactly, making the analysis easier to understand and to compare from different days.
+\subsection{Adding country information}
 
-\subsubsection{Normalized Frequency}
-We compute the number of the messages of each time-window by signal. To normalize frequency, we compute \textit{z-score} as following:
+\subsubsection{Adding primary languages}
 
-\begin{equation} \label{eq:1}
-zscore = \frac{x_{i} - \mu_{k} }{\sigma_{k}}
-\end{equation}
+\subsection{Merging Labels}
 
-where $x_{i}$ is the frequency of the current $i$ time-window, $\mu_{k}$ and  $\sigma_{k}$ are mean and standard deviation of the previous $k$ time-windows respectively.
+\subsection{Adding Crisis Categories}
 
-%\begin{itemize}
-%	\item We calculate the number of messages of each time windows for each signal.
-%	\item In our first approach, we compute max-min normalization and z-score normalization of all time windows for each signal.
-This approach is wrong because we do not know the next values in the streaming.
-We just know the previous values.
-%	\item We compute the z-score normalization using past values for each signal independently.
-We do not use rolling/moving z-score because some signals have Null values of frequency.
-It is happen in smallest cities or states.
-%\end{itemize}
-\subsubsection{Interarrival Time}
+\section{Exploratory Analysis}
 
-To characterize the urgency of the messages during a time-window, we use our previous results \cite{DBLP:conf/fdia/Sarmiento17} to compute the \textit{interarrival time} which is defined as $d_{i} = t_{i+1} - t_{i}$, where $d_{i}$ denotes the difference between two consecutive social media messages $i$ and $i+1$ that arrived in moments $t_{i}$ and $t_{i+1}$ respectively.
-Using this definition and according the work of \citeauthor{kalyanam2016prediction} \cite{kalyanam2016prediction}, high-activity events have a high-frequency in the first bins represented by values $d_{i} \approx 0$.
+\subsection{Computing Features}
 
-To quantify a high-frequency in very small values of $d_{i}$, we compute the measures \textit{skewness} and \textit{kurtosis}, which represent the asymmetry and the tailedness of the shape of probability distribution respectively \cite{mardia1970measures}.
-Finally, we apply the equation \ref{eq:1} over \textit{skewness} and \textit{kurtosis} to calculate variation based on previous values.
+\subsubsection{Content Features}
 
-%\begin{itemize}
-%	\item Inter-arrival time represent the difference between 2 consecutive messages for one signal.
-In this case, a small difference between a set of messages represent the urgency or relevancy in a specific time (reference to plosone paper)
-%	\item First, we compute inter-arrival time for each time windows independently.
-We summary results using quantiles, mean, min and max values.
-The problem with this approach is that we can not compare values between signals because each signal has different frequencies and behaviours along the time.
-%	\item We compute skewness and kurtosis which represent a measure of the asymetry and "taildness" of the probability distribution respectively.
-In both case, we calculate the measures using previous values such as z-score normalization.
-%\end{itemize}
+\subsubsection{Grammatical Features}
 
-%\subsection{Subsetting Signals}
-%\begin{itemize}
-%	\item Several signals have a null values along the time.
-For example, any city that is not a big city or capital city, does not a continuous frequency.
-To avoid this behaviour, we use cities just for compute the state frequency signal and later we discard them.
-Finally, we just monitor states and country hierarchy.
-%	\item Furthermore, we remove gps signal because just 1-3 percent of messages have gps activated.
-%\end{itemize}
+\subsubsection{Entities-Based Features}
 
-\subsection{Geographic Spread}\label{sssec:geospread}
+\subsubsection{Emotional Features}
 
-An emergency situation that affects and mobilizes response in a small area is defined as \textit{focalized}, while a disaster with a large geographic impact is defined as \textit{diffused} \cite{olteanu2015expect}.
-Using this definition, we extend this concept to represent neighborhoods between locations obtained from section \ref{sssec:geohie}.
-For that purpose, we create an \textit{adjancency matrix} $M$, where $M_{i,j} = 1$ represents if two locations are geographically connected and $M_{i,j} = 0$ if they are not connected.
-For instance, if an event is diffused (e.g., earthquake), the detection should be in adjacent-locations independentely of metadata-level.
-On the other hand, if an event is focalized (e.g., terrorist attack), just one location should be detected but in different metadata-levels simultaneously.
+\subsection{Hazard Categories Analysis}
 
-\begin{table}
-	\caption{List of earthquakes studied as ground truth, sorted by date.}
-	\label{tab:eqs}
-	\begin{tabular}{lccl}
-		\toprule
-		Country&Datetime (UTC)&Magnitude (Mw)& Language\\
-		\midrule
-		Italy & 2016-10-26 17:10:36 & 5.5 & Italian \\
-		Italy & 2016-10-30 06:40:17 & 6.6 & Italian\\
-		Chile & 2016-12-25 14:22:26 & 7.6 & Spanish\\
-		Chile & 2017-04-23 02:36:06 & 5.9 & Spanish\\
-		Chile & 2017-04-24 21:38:28 & 6.9 & Spanish\\
-		\midrule
-		%\bottomrule
-	\end{tabular}
-\end{table}
+\subsection{Hazard Sub Categories Analysis}
 
-\section{Dataset}
+\subsection{Hazard Types Analysis}
 
-Data is collected from Twitter Public Streaming API, which allows access to subsets equal to $1\%$ of public status descriptions in real-time.
-With this tool, we can retrieve either messages using a set of keywords or messages from specific locations setting a bounding box.
-In our approach, we get entire subsets of messages without use keywords or specific locations.
-Then, we retrieve random messages about any topic and any place in the world.
+\begin{itemize}
+	\item $82\%$ english data and $8\%$ spanish and italian each.
+	\item Most events ocurred in United States
+\end{itemize}
 
-\subsection{Ground Truth}
-
-We analyse five earthquakes with magnitudes between $5.5Mw$ and $7.6Mw$\footnote{Mw: the moment magnitude scale}, ocurring in Italian-speaking and Spanish-speaking countries between October 2016 and April 2017 (Table \ref{tab:eqs}).
-For that purpose, we collect $20$ million of messages 12-hours before and after the emergency situation events.
-
-According our proposal we create both the gazetteer hierarchies\footnote{http://users.dcc.uchile.cl/$\sim$hsarmien/gazetteer.html} for each country and construct the signals based on each hierarchy and metadata-level.
-As a result of the number of messages of each signal (Table \ref{tab:message_signal}), we discard all signals related to city hierarchy since a great amount of small cities have zero frequency in a normal situation unlike to capital or metropolitan cities.
-
-\begin{table}
-	\caption{Number of messages by signal.}
-	\label{tab:message_signal}
-	\begin{tabular}{ccl}
-		\toprule
-		Hierarchy&Metadata-level&Messages\\
-		\midrule
-		All & All & 87,291 \\
-		\midrule
-	    \multirow{3}{*}{Country} & Tweet Text & 11,584 \\
-	    & User Location & 25,313\\
-	    & Tweet Text - User Location & 1,417\\	
-		\bottomrule
-		\multirow{3}{*}{State} & Tweet Text & 4,110 \\
-		& User Location & 13,352\\
-		& Tweet Text - User Location & 86\\
-		\midrule
-		\multirow{3}{*}{City} & Tweet Text & 1,415 \\
-		& User Location & 8,971\\
-		& Tweet Text - User Location & 20\\
-		\bottomrule
-	\end{tabular}
-\end{table}
-
-\subsubsection{Labeled Emergency Situations}
-
-The exactly datetime event is obtained from National Seismology Agency in Chile\footnote{http://www.sismologia.cl/} and 
-National Institute of Geophysics and Volcanology in Italy\footnote{http://www.ingv.it/it/}.
-With the purpose of labeling a time-window as positive class (detection), we set as detection those time-windows with positive variation in frequency, skewness and kurtosis with respect to the normalization of the previous values.
-Moreover and according to (Figure \ref{fig:labeled}), we include the three next time-windows after the event to compensate the imbalance between classes given that after these number of time-windows, the variation in the features decrease.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{img/labeled2.png}
-	\caption{Average variation in emergency situations between time-windows.}
-	\label{fig:labeled}
-\end{figure}
-
-\section{Experiments}
-
-Our filtering task can be seen as binary classification task.
-The positive class (\textit{detection label}) corresponds to messages related to instantaneous emergency situations, while the negative class (\textit{nothing label}) corresponds to the remaining or non-related to crisis situations.
-
-To classify messages we employed traditional binary classifier Support Vector Machine \textit{(SVM)}.
-As a result of the analysed data scattering (Figure \ref{fig:scatter}), we separated country and state in different datasets and set both kernels and classification parameters independentely.
-On the one hand, \textit{country} classifier uses a polynomial kernel and strict-parameters for gamma, cost and weights since a great amount of messages are included in country hierarchy as an effect of the minor hierarchies.
-On the other hand, \textit{state/region} classifier uses a linear kernel with soft-weights and cost.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{img/scatter.png}
-	\caption{Relationship between features in country and state hierarchy. Red circles represent positive class (\textit{detection}) and blue circles represent negative class (\textit{nothing}).}
-	\label{fig:scatter}
-\end{figure}
-
-Given that an emergency situation is not an usual event, we have an highly unbalanced data respect to the classes after labeled ($1\approx2\%$ of positive class corresponding to \textit{detection}).
-Therefore, we used \textit{under-sampling} \cite{lunardon2014rose} over country and state datasets increasing our possitive class to $15\approx18\%$.
-Additionally, to validate our model, we used \textit{5-fold cross-validation} where one earthquake dataset is used as testing and the remaining earthquakes dataset as training.
-
-Table \ref{tab:perfomance} shows the average results of our model applying 5-fold cross validation.
-In order provide a extended analysis about incorrect labels and time-windows, we include the metric \textit{False Positive Rate (FPR).}
-
-\subsubsection{Independent Analysis of Hierarchies} 
-Our first analysis is just considering the hierarchies as isolated detections.
-The top of the Table \ref{tab:perfomance} shows the results considers only the prediction over each label in our datasets.
-As noted above, the assignation from the lowest level (\textit{city}) to the highest (\textit{country}) in the gazetteer hierarchy generated high frequency of messages which cause multiple \textit{bursts} in our country signal for non  emergency situations.
-This concept can explain the values of Precision (\textit{P}) and \textit{FPR}.
-
-In addition to the analysis of number of detections by labels, we also study the number of detections by time-windows.
-For this analysis, we search the time-windows for each hierarchy where the all metadata-levels are well classified with correct class.
-According to the results shown on the middle of the Table \ref{tab:perfomance}, when we analyse country and state independentely the values of Precision, F1 and FPR have worst values than the analysis by label .
-\begin{table}
-	\caption{Average performance of 5-fold cross-validation by hierarchy and geographic spread (G.S.)}.
-	\label{tab:perfomance}
-	\centering
-	\begin{tabular}{c|lcccc}
-		\toprule
-		&Hierarchy &P &R &F1 &FPR\\
-		\midrule
-		\parbox[t]{2mm}{\vspace{-0.2cm}\multirow{3}{*}{\rotatebox[origin=c]{90}{\small{label}}}} 
-		& Country & 0.3 & 0.83 & 0.45& 0.14\\		
-		& State &0.35& 0.83& 0.5 & 0.08\\
-		\midrule
-		\parbox[t]{2mm}{\vspace{0cm}\multirow{5}{*}{\rotatebox[origin=c]{90}{\small{time-window}}}}
-		& Country & 0.15& 0.77 & 0.25 & 0.15\\
-		& State & 0.17 & 0.88 & 0.29& 0.12\\
-		& Country-State& 0.35 & 0.7 & 0.47 & 0.03\\
-		& Country(2)-State with G.S. & 1 & 0.64 & 0.78 & 0 \\
-		& Country(3)-State with G.S. & 1 & 0.47 & 0.64 & 0\\
-		\bottomrule
-	\end{tabular}
-\end{table}
-
-\subsubsection{Dependent Analysis of Hierarchies}
-Our second analysis considered the hierarchies as non-isolated detections.
-In the results explained above, we considered country and state hierarchy independently, which is not a correct analysis because an emergency situation affects states and country at the same time.
-For this reason, we inspected the time-windows where all metadata-level for country and state hierarchy have a correct detection simultaneously.
-The results are shown in the row \textit{Country-State} in the Table \ref{tab:perfomance}.
-In contrast to the independent analysis of country and state, we improved the Precision, F1 and FPR values as a consequence of a smaller amount of the time-windows related to non emergency situations are assigned as detection.
-However, when we see the value obtained for FPR ($FPR = 0.03$), this rate represents an incorrect number of time-windows assigned as detection equal to $23$.
-This means that we have $23$ new emergency situations detected by our classifier.
-
-\subsubsection{Geopraphic Spread Analysis}\label{sssec:geospreadanalysis}
-Our third analysis considered the hierarchies as non-isolated detections and applies the Geographic Spread (G.S.).
-Using the \textit{Adjacency Matrix} to represent neighborhoods between regions/states, we considered as a correct detection those time-windows where the state/s classified as detection are defined as \textit{Focalized} or \textit{Diffused} and exist dependency between hierarchies.
-
-\begin{table*}
-	\caption{Online evaluation by time-windows (T-W) using Country(2)-State with G.S. method}
-	\label{tab:online1}
-	\begin{tabular}{lccccp{3.5cm}c}
-		\toprule
-		Event &Detected T-W & T-W Before Event & T-W After Event & Delay (min) & Top 3 Bigrams\\
-		\midrule
-		Premier League Soccer Matches & 2& - & - & - & \small{(man, utd), (new, year), (happy, new)} \\
-		Westminster Terrorist Attack& 13 & 0 &13 & 32& \small{(stay, safe), (terror, attack), (safe, everyone)}\\
-		Manchester Terrorist Attack& 12& 1& 11& 23& \small{(ariana, grande), (incident, arena), (grande, concert)}\\
-		London Terrorist Attack & 14 & 7 & 7 & 36 & \small{(stay, safe), (incident, bridge), (borough, market)}\\
-		U.K. Elections& 5 & - & - & - & \small{(theresa, may), (vote, labour), (van, dijk)}\\
-		Adele Live in Wembley& 9& 7&2&-& \small{(elland, road), (new, times), (phil, jackson)}\\
-		England vs Slovenia Soccer Match & 4& 4 & 0 & - & \small{(simon, brodkin), (join, us), (theresa, may)}\\
-		Metallica Live in London& 4 & 4 & 0 & - & \small{(always, said), (chance, win), (carabao, cup)}\\
-		\bottomrule
-	\end{tabular}
-\end{table*}
-
-\begin{table*}
-	\caption{Online evaluation by time-windows (T-W) using Country(3)-State with G.S. method}
-	\label{tab:online2}
-	\begin{tabular}{lccccp{3.5cm}c}
-		\toprule
-		Event &Detected T-W & T-W Before Event & T-W After Event & Delay (min) & Top 3 Bigrams\\
-		\midrule
-		Premier League Soccer Matches & 0& - & - & - & \hfill \break \\
-		Westminster Terrorist Attack& 4 & 0 &4 & 32& \small{(terror, attack), (stay, safe), (terrorist, attack)}\\
-		Manchester Terrorist Attack& 2& 0& 2& 23& \small{(ariana, grande), (praying, everyone), (everyone, affected)}\\
-		London Terrorist Attack & 1 & 1 & 0 & - & \small{(ariana, grande), (around, world), (lady, gaga)} \\
-		U.K. Elections& 0 & - & - & - & \hfill \break\\
-		Adele Live in Wembley& 0& 0&0&-& \hfill \break\\
-		England vs Slovenia Soccer Match & 1& 1 & 0 & - & \small{(per, day), (menswear, sample), (closed, roads)}\\
-		Metallica Live in London& 2 & 2 & 0 & - & \small{(happy, birthday), (chance, win), (always, said)}\\
-		\bottomrule
-	\end{tabular}
-\end{table*}
-
-In addition to the results of the dependency analysis explained above, we see that a large amount of time-windows for country hierarchy ($\approx 82\%$) have more than one metadata-level when exist a correct detection.
-This can be explained since an emergency situation produce a collective reaction on the level of body of the message (\textit{tweet text}), users sharing any messages with profile location in a specific country (\textit{user location}) or mixing both concepts (\textit{tweet text - user location}).
-
-Considering the geographic spread by states and the number of metadata-levels by country hierarchy, we analysed the results shown on the bottom of the Table \ref{tab:perfomance}.
-On the one hand, the row with value equal to \textit{Country(2)-State with G.S.} represents the detection when we considered at least two metadata-levels for the country hierarchy and the geographic spread for states.
-In contrast to the the previous analyses, we improved the values of the Precision, F1 and FPR.
-The last metrics are very important because there are no time-windows incorrectly assigned as emergency situations.
-Consequently, the Recall values decrease which means that our method remove some time-windows classified as detection.
-Beside the percent of emergency situations detected is equal to $100\%$ with a average delay equal to $10.4$ minutes ($min=6$, $max=14$) from the impact of the event to the first detection.
-
-On the other hand, \textit{Country(3)-State with G.S.} represent the detection when we considered three metadata-levels for country hierarchy and the geographic spread for states.
-Similar to  \textit{Country(2)-State with G.S.}, we improved the values of Precision, F1 and FPR but our recall decrease from $R = 0.58$ to $R = 0.47$, detecting  $80\%$ of the emergency situations with a average delay equal to $11.5$ minutes ($min=8$, $max=14$) from the impact of the event to the first detection.
-
-\subsection{Online Evaluation}
-For our evaluation in the Twitter Public Stream, we trained classifier with five earthquakes identified in our ground truth.
-Furthermore, our evaluation dataset is formed by eight different events that occurred in England between December 2016 and October 2017.
-For each event we considered the full-day in which they ocurred.
-The main goals of this evaluation is to know the capacity of our method to detect emergency situations and discard those non-related to emergency events that involve location references.
-Geographic spread analysis is used to evaluate our method because decrease the number of false positives detection.
-In the same way of the experiments in Section \ref{sssec:geospreadanalysis}, we compare the results using the two presented methods respect to the number of metadata-levels by country hierarchy.
-
-As can be noted on Table \ref{tab:online1} and Table \ref{tab:online2}, we study three terrorist attacks and five high-impact real-world events related to soccer matches, music concerts and political elections.
-In the case of Premier League Soccer Matches and U.K Elections, we can not identify the beginning of the event, since in the first one there are many soccer matches during the analysed day and in the second one there is no a specific start time.
-In order to know the topics when our method detects an event, we computed the Top 3 Bigrams in the detected time-windows.
-Also, we calculated the delay time just for emergency events.
-
-On the one hand, the first evaluation \textit{Country(2)-State with G.S.} has full detection of the terrorist attacks with average delay time equal to $30.3$ minutes.
-These detections are related to the event given that the bigrams represent terms associated with crisis situations.
-However, the London Terrorist Attack has $50\%$ of the detected time-windows after the event, which means that there are seven time-windows non-related to emergency situations.
-Besides the crisis situations analysis, we also study the number of detected time-windows in non-related to emergency situation events.
-In the same way, we have a large amount of misclassified time-windows that do not represent crisis situations as we can see in the Top 3 Bigrams for each non-related to event.
-
-On the other hand, the second evaluation \textit{Country(3)-State with G.S.} decreases the number non-related to emergency situations events detected as crisis situations.
-We can see three time-windows in two events detected as emergency situations (England vs Eslovenia, and Metallica Live in London).
-In these cases, the time-windows are detected before the event and corresponding a non emergency situations according to the bigrams.
-Furthermore, when we analysed the number of the detected emergency situations, two-thirds ($66\%$) of the events are detected correctly with average delay time equal to $30.3$ minutes.
-In the case of London Terrorist Attack, our method detects one time-window before the event but the bigrams describe that the detections do not correspond to crisis situations.
-
-%Unlike to  \textit{Country(2)-State with G.S} method, \textit{Country(3)-State with G.S} method decreases the number of detection by time-windows given that latter is stricter than the first one respect to the number of country detection by metadata-level.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{img/delay.png}
-	\caption{Delay time and number of locations in the first detection for diffused and focalized emergency situations.}
-	\label{fig:delay}
-\end{figure}
-
-\section{Discussion}
-
-Our findings suggest that there is evidence to detect an emergency situation based on anomaly frequency of messages that contain locations for a specific country.
-Indeed, our method based on the number of metadata-levels by country hierarchy and geographic spread by state, detects a $80\%$ of the events related to emergency situations as we could demonstrate in our ground truth.
-Also, our method is independent of the textual features because we apply the model over different languages as Spanish, Italian and English.
-Furthermore, we test our model in different types of events such as earthquakes and terrorist attacks, and also on different magnitudes (in the case of earthquakes) and number of affected people (e.g., Manchester Terrorist Attack vs Westminster Terrorist Attack).
-
-However, when we apply our method in the on-line evaluation, we detect $66\%$ of the emergency situations that affected England.
-This explains that the signals, and for various reasons: the number of active users in United Kingdom\footnote{http://www.statista.com/statistics/242606/number-of-active-twitter-users-in-selected-countries/ visited on January 2018} which can affect the anomaly frequency of the messages since there exists a high daily average activity of the messages; similar locations in other countries (York $\approx$ New York); and soccer teams with names of cities (Manchester United, Liverpool).
-These issues also can affect the number of false positive detection in which in the case of England was $30\%$ of the non-related to emergency events.
-
-Regarding the geographic spread where we define an emergency situation as diffused or focalized, we find some evidence that differentiates them.
-In the case of diffused events, the delay time of the our first detection was less than $12$ minutes and in focalized events was greater than $30$ minutes (Figure \ref{fig:delay}).
-This explains that, in diffused events such as earthquakes, a high number of people are affected (thousands or millions) at the same time by an event which generates a collective reaction in social media in the locations where the event impacted.
-In Figure \ref{fig:delay}, we can see that earthquakes have at least two detected locations in the first detection (except Italy EQ2).
-In contrast, focalized events have less amount of eyewitness (hundreds or thousands) then when the users share messages in social media, the frequency does not affect the average daily message of the country in the first minutes.
-This can be explained in Figure \ref{fig:delay} where the terrorist attacks have just 1 detected location in the first detection.
-
-Additionally, the delay time can be different for many reasons: datetime of the event (for example, during the early hours), few differences with the end of the current time-window, type of the affected locations (rural, urban cities) and the number of active users by locations.
-
-%\begin{itemize}
-%	\item explain that we dont need compute textual features as bigrams, top-k terms.
-%	\item explain that we just need de gazetteer tree
-%	\item explain when our methodology could be not working. eg. why not working correctly in england. country with many language. How event like soccer match are filtered and when not working
-%	\item Delay time: explain difference between focalized and diffused event. People feel the event at the same time, activity in affected locations.
-%	\item Number of locations affected in the first and second detection for focalized and diffused (plot)
-%	\item Diffused: mercalli and locations detected.
-
-%\end{itemize}
-
-\section{Conclusion}
-
-In this paper we have presented a methodology for detecting an emergency situation based on location for a specific country.
-This approach is independent of the textual features and can be used in different types of events and languages.
-We show that the users act as self-organized in the affected locations like citizen sensors when an emergency situation occurs.
-We furthermore have presented an analysis of geographic spread for different types of events that can be categorized.
-However, our experiment considers just a small portion of emergency situations, which is not representive for all types of crisis situations according to either the hazard type (natural or human-induced), temporal development (instantaneous or progressive) or geographic spread (diffused or focalized).
-
-There are many things that can improve our results.
-We will add Point of Interest to our gazetteer tree to increase the frequency by time-windows in each hierarchy.
-Furthermore, we will add more non-textual features as number of retweets and tweets, unique locations detected and special locations.
-We also plan to study the relevance of the different metadata-levels and assign weights for each.
-Finally, we will create a web application to visualize events in real-time.
-
-%\begin{acks}
-
-%\end{acks}
+\begin{acks}
+This work has been partially funded by the Millennium Institute for Foundational Research on Data.	
+\end{acks}
 
 \bibliographystyle{ACM-Reference-Format}
+\balance
 \bibliography{bio}
